---
title: "Figure Code"
output: html_notebook
---

```{r}
library(DESeq2)
library(tidyverse)
library(data.table)
library(biomaRt)
library(limma)
library(ggrepel)
ensembl <- useEnsembl(biomart = "genes", dataset = "mmusculus_gene_ensembl")
bmart1<-getBM(attributes = c("ensembl_gene_id",
                               "external_gene_name","chromosome_name","gene_biotype"), mart = ensembl)
bfilt<-bmart1%>%filter(chromosome_name%in%c(as.character(1:19),"X","Y"),gene_biotype%in%c("protein_coding"))
```

Developmental ATAC-seq Motif enrichment Plots
```{r}

```


Layer/Age KLF plot
```{r}
quickTPM_line<-function(TPMmat,Genes=NULL,Pattern=NULL,wrap=FALSE,logscale=F,lwd=1,L4s1=TRUE){
  if (is.null(Genes)&&is.null(Pattern)){
    stop("Please Select at Least 1 Gene")
  }
  if(!any(names(TPMmat)=="external_gene_name")){names(TPMmat)[which(grepl(names(TPMmat),pattern="gene",ignore.case = T))]="external_gene_name"}
  if (!is.null(Genes)&&is.null(Pattern)){
    TPMmut <-TPMmat%>%filter(external_gene_name%in%Genes)
  }
  if (is.null(Genes)&&!is.null(Pattern)){
    TPMmut<-TPMmat%>%filter(grepl(external_gene_name,pattern=Pattern,ignore.case = T))
  }
    if(L4s1){TPMmut<-TPMmut%>%dplyr::select(matches("p2|p30|L6|gene"))}else{TPMmut<-TPMmut%>%dplyr::select(-matches("p7|p30|p3"))}
    TPMmut<-TPMmut%>%pivot_longer(-external_gene_name)%>%
      mutate(CellType=str_extract(name,pattern = "L[4-6]+"),Age=ifelse(grepl(name,pattern = "p2"),"p2","p30"))%>%
      mutate(Age=factor(Age,levels = c("p2","p30")))
    if(logscale){TPMmut<-TPMmut%>%mutate(value=log2(value+1))}
    TPMmut<-TPMmut%>%
      group_by(CellType,Age,external_gene_name)%>%mutate(Groupmean=mean(value),Groupsd=sd(value))
    TPMplot<-TPMmut %>% ggplot(aes(x=Age,y=Groupmean,group=interaction(CellType,external_gene_name),col=external_gene_name,linetype=CellType))+
      geom_line(lwd=lwd)+geom_errorbar(lwd=lwd,aes(ymin=Groupmean-Groupsd,ymax=Groupmean+Groupsd),width=0.05)+
      ylab("Normalized Expression (TPM)")+xlab("Age")+labs(linetype="Cell Type",col="Gene")+scale_x_discrete(expand=c(0.1,0.1))+
      theme_classic()
    if(wrap){TPMplot=TPMplot+facet_wrap(. ~ external_gene_name, scales="free")}
    if(logscale){TPMplot<-TPMplot+labs(y="Normalized Counts (log2TPM)")}
    return(TPMplot)
}

L46_TPM<-read.csv("~/Documents/Counts/p2p30/p2p30_mm10_TPM.csv")

plotout<-quickTPM_line(L46_TPM,Genes=c("Klf9","Klf7","Klf13","Klf6"),L4s1 = T)+
  scale_color_manual(values=c("chartreuse3","chartreuse4","tomato","red3"),breaks=c("Klf6","Klf7","Klf13","Klf9"))+
  facet_grid(~factor(external_gene_name,levels=c("Klf6","Klf7","Klf13","Klf9")))+
  theme(text = element_text(size = 18),strip.background = element_blank())+
  scale_linetype_discrete(name="Cell Type",labels =  c("Layer IV (Rorb+)","Layer VI (56L)"))
```


L4 p2/7/30 Klf plot
```{r}
quickTPML4<-function(TPMmat,Genes=NULL,Pattern=NULL,wrap=FALSE,LinePlot=FALSE,logscale=F){
  if (is.null(Genes)&&is.null(Pattern)){
    stop("Please Select at Least 1 Gene")
  }
  if (!is.null(Genes)&&is.null(Pattern)){
    TPMmut <-TPMmat%>%filter(external_gene_name%in%Genes)
  }
  if (is.null(Genes)&&!is.null(Pattern)){
    TPMmut<-TPMmat%>%filter(grepl(external_gene_name,pattern=Pattern,ignore.case = T))
  }
    TPMmut<-TPMmut%>%pivot_longer(-external_gene_name)%>%
      mutate(Age=str_extract(name,pattern="p[0-9]+"))%>%
      mutate(Age=fct_relevel(Age,"p2","p7","p30"))
    if(logscale){TPMmut<-TPMmut%>%mutate(value=log2(value+1))}
    TPMmut<-TPMmut%>%mutate(Group=Age)%>%
      group_by(Group,external_gene_name)%>%mutate(Groupmean=mean(value),Groupsd=sd(value))
  if (LinePlot==F) {
      TPMplot<-TPMmut %>% ggplot(aes(x=external_gene_name,y = Groupmean,fill=Group)) + geom_col(col="black",position="dodge") +
      geom_point(position=position_dodge(width=0.9))+
      geom_errorbar(aes(ymin=Groupmean-Groupsd,ymax=Groupmean+Groupsd,x=external_gene_name),position =position_dodge(width=0.9),width=0.2) +
      ylab("Normalized Expression")+xlab("Gene")+theme_classic()+theme(legend.position = "bottom")+
        theme(axis.text.x = element_text(angle = 45,size = 12,vjust=0.5))
    } else if(LinePlot==T){
        TPMplot<-TPMmut%>%ggplot(aes(x=Group,y=Groupmean,group=external_gene_name,color=external_gene_name))+
    geom_line(linewidth=1)+
    geom_errorbar(aes(ymin=Groupmean-Groupsd,ymax=Groupmean+Groupsd),width=0.2)+
    labs(y="Normalized Expression (TPM)",x="Age",col="Gene")+
    theme_classic()
    }
    if(wrap){TPMplot<-TPMplot+facet_wrap(~external_gene_name,scales="free")}
    if(logscale){TPMplot<-TPMplot+labs(y="Normalized Expression (log2TPM)")}
    return(TPMplot)
}

L4TPM<-read.csv("~/Documents/Counts/p2p30/p2p30_L4_RefGenes.csv",row.names = 1)

outplot<-quickTPML4(L4TPM,Genes=c("Klf9","Klf7","Klf13","Klf6"),LinePlot = TRUE)+
  scale_color_manual(values=c("chartreuse3","chartreuse4","tomato","red3"),breaks=c("Klf6","Klf7","Klf13","Klf9"))+
  theme(text = element_text(size = 18),strip.background = element_blank())
```


Arlotta paper development plot
```{r}
quickplotArlotta<-function(TPMmat,Genes=NULL,Pattern=NULL){
    if (is.null(Genes)&&is.null(Pattern)){
    stop("Please Select at Least 1 Gene")
  }
  if (!is.null(Genes)&&is.null(Pattern)){
    TPMmut <-TPMmat%>%filter(GeneName%in%Genes)
  }
  if (is.null(Genes)&&!is.null(Pattern)){
    TPMmut<-TPMmat%>%filter(grepl(GeneName,pattern=Pattern,ignore.case = T))
  }
  TPMmut%>%pivot_longer(cols = -GeneName)%>%mutate(celltype=str_extract(name,"Tle4|Cux2"),Age=str_split_fixed(name,"_",n=2)[,1])%>%
    group_by(GeneName,celltype,Age)%>%
    mutate(mean=mean(value),Age=factor(Age,levels = c("E18","P1","P3","P7","P21","P48")))%>%
    ggplot(aes(x=Age,y=mean))+
    geom_line(aes(group=interaction(celltype,GeneName),col=GeneName,linetype=celltype),linewidth=1.2)+
    geom_point(aes(y=value,col=GeneName))+
    labs(y="Normalized Counts (FPKM)",x="Age",col="Gene")+
    theme_classic()
}

ArlottaTPM<-read.csv("~/Documents/Counts/Arlotta_devFPKM.csv",row.names = 1)

plot1<-quickplotArlotta(ArlottaTPM,Genes=c("Klf9","Klf7"))+scale_linetype(name="Cell type")+
  theme(text =element_text(size=18),legend.position = "right",legend.text = element_text(size=12),legend.title = element_text(size=16))+
  scale_color_manual(values=c("chartreuse4","red3"))+
    scale_linetype_discrete(name="Cell Type",labels =  c("Layer II/III (Cux2+)","Layer VI (Tle4+)"))
```

Layer conservation plot
```{r}
GeneCounts<-read.csv("~/Documents/ATACseq_Project/FinalCounts/InternalRNA_logCPM_byGene_batchcor.csv",row.names = 1)
bmart1<-read.csv("~/Documents/ATACseq_Project/RegressionStuff/ENS_info2020.csv",row.names = 1)

PeakNames<-"_L[1-6]+_"

GeneCountsfilt=GeneCounts%>%dplyr::select(matches(PeakNames))%>%
  dplyr::select(-matches("_p[1-9]_"))

Cellavg<-GeneCountsfilt%>%rownames_to_column("X")%>%gather("Sample","Expression",-X)%>%
  mutate(CellType=str_split_fixed(Sample,pattern = "_combo|_[0-9]$",n=2)[,1])%>%
  mutate(CellType=str_split_fixed(CellType,pattern = "_",n=2)[,2])%>%
  group_by(CellType,X)%>%
  dplyr::summarize(avg=mean(Expression))%>%spread(CellType,avg)%>%
  left_join(bmart1,by=c("X"="ensembl_gene_id"))%>%dplyr::select(-X)

names(Cellavg)[1:6]<-c("Layer 2/3","Layer 4","Layer 5","Layer 5 (2)","Layer 6","Layer 6 (2)")
  
Klfs<-Cellavg%>%filter(grepl(external_gene_name,pattern="^Klf[0-9]+"))%>%column_to_rownames("external_gene_name")

out<-pheatmap(Klfs,cluster_cols = F,border_color = NA,angle_col = 315,fontsize_col = 12,fontsize_row = 12,clustering_method = "ward.D")


```

CRISPRi test plot
```{r}
x<-read.csv("~/Documents/CRISPRi/qPCR/CRISPRi_summary.csv")
names(x)<-c("FC","Target","Group","Sample","Plate")
x<-x%>%dplyr::mutate(Group=ifelse(Group=="GFP","Scramble gRNA AAV","Triple KD AAV"))%>%dplyr::group_by(Group,Target)%>%dplyr::mutate(fcmean=mean(FC))%>%arrange(Target,Group)
n<-length(unique(x$Sample))
n2<-length(unique(x$Target))
x$controlmean<-rep(x$fcmean[seq(1,n*3,n)],each=n)
x<-x%>%dplyr::mutate(fcnorm=FC/controlmean)%>%dplyr::mutate(normmean=mean(fcnorm),fcsd=sd(fcnorm))
out<-ggplot(x)+geom_col(aes(x=Target,y=normmean,fill=Group),col="black",position="dodge")+
  geom_point(aes(x=Target,y=FC,fill=Group),position = position_dodge(width=0.9),show.legend = F)+
  geom_errorbar(aes(ymin=normmean-fcsd,ymax=normmean+fcsd,x=Target,fill=Group),position =position_dodge(width=0.9),width=0.2)+
  scale_y_continuous(expand=c(0,0),limit=c(0,ceiling(max(x$FC))))+theme_classic()+
  scale_fill_brewer(palette = "Set1")+
  labs(x="Measured Gene",y="Relative mRNA Levels",fill="gRNA Target")+theme(legend.position = "bottom")+geom_hline(yintercept = 1,linetype="dashed")+
  theme(text=element_text(size=16),legend.position = "right",legend.text = element_text(size=12))
```

KLF qPCR Plot
```{r}
KLF<-read.csv("~/Documents/CRISPRi/qPCR/KLF9_13_913_mCherry.csv")
#KLF<-gather(KLF, "Target", "FC", KLF13, KLF9)
KLF$Group[KLF$Group=="ScrR"]="Scramble gRNA"
KLF<-KLF%>%group_by(Target,Group)%>%mutate(fcmean=mean(FC))%>%arrange(Target,desc(Group))%>%
  mutate(Group=factor(Group,levels = c("Scramble gRNA","Klf13","Klf9","Klf9/13")))
n<-length(unique(KLF$Sample))
KLF$controlmean<-c(rep(KLF$fcmean[1],n),rep(KLF$fcmean[n+1],n))
KLF<-KLF%>%mutate(fcnorm=FC/controlmean)%>%mutate(normmean=mean(fcnorm),fcsd=sd(fcnorm))
y<-ggplot(KLF)+geom_col(aes(x=Target,y=normmean,fill=Group),col="black",position="dodge")+geom_point(aes(x=Target,y=FC,fill=Group),position = position_dodge(width=0.9),show.legend = F)+
  geom_errorbar(aes(ymin=normmean-fcsd,ymax=normmean+fcsd,x=Target,fill=Group),position = position_dodge(width=0.9),width=0.2)+scale_y_continuous(expand=c(0,0),limit=c(0,4))+theme_classic()+
  labs(x="Measured Gene",y="Relative mRNA Levels",fill="gRNA Target")+
  theme(legend.position = "right",text=element_text(size=16))+geom_hline(yintercept = 1,linetype="dashed")+
  scale_fill_manual(values=c("blue4","sandybrown","salmon2","firebrick2"))
```
All CRISPRi summary
```{r}
infile<-fread("~/Documents/CRISPRi/qPCR/CRSIsummary.csv")
tempfile<-infile%>%dplyr::select(Target,SampID,KD)%>%
  group_by(Target)%>%mutate(KD=KD*100)%>%mutate(kdmean=mean(KD),kdsd=sd(KD))

y<-ggplot(tempfile,aes(x=Target,y=kdmean,fill=Target))+geom_bar(stat="summary",position="dodge",color="black",linewidth=0.75)+
  geom_errorbar(aes(ymin=kdmean-kdsd,ymax=kdmean+kdsd,x=Target),position = position_dodge(width=0.9),width=0.2)+
  geom_point(aes(x=Target,y=KD),position = position_dodge(width=0.9),show.legend = F)+
  scale_y_continuous(expand=c(0,0))+theme_classic()+scale_fill_brewer(palette = "Set1")+
  labs(x="gRNA Target",y="Knockdown %")+theme(legend.position = "none")+geom_hline(yintercept = 100,linetype="dashed")
y+theme(axis.title =element_text(size=15),axis.text.x =element_text(size=15,angle = 45,vjust = 0.6),axis.text.y=element_text(size=17))
```

PCA plot
```{r}
#Using DESEQ2 results from Klf9/13 as input
library(limma)
load("~/Documents/CRISPRi/SeqAnalysis/Klf913old_DESeq2.RDa")
rlg <- rlog(ddsk913, blind=FALSE)

mat<-removeBatchEffect(assay(rlg),rlg$Batch)
rlg_batch<-rlg
assay(rlg_batch)<-mat
tst<-plotPCA(rlg_batch,intgroup=c("Target"),ntop=500)+theme_bw()+labs(color= "Target")
tst<-tst+scale_color_manual(values=c("blue4","sandybrown","salmon2","firebrick2"),name="gRNA Target",labels=c("Scramble gRNA","Klf13","Klf9","Klf9/13"))+
  theme(axis.title =element_text(size=14),legend.position = "bottom")+
  geom_point(color="black",pch=21,size=4)+geom_point(aes(fill=group),size=3.5,show.legend = F)
```

Get DEGs from KLF913 data
```{r}
lfc<-log2(2)
pval<-0.05
KLF913old<-lfcShrink(ddsk913, coef="Target_Klf913_vs_ScrR",type="apeglm")

K913df_old<-as.data.frame(KLF913old)%>%rownames_to_column("ENSID")%>%
  left_join(bmart1,by = c("ENSID"="ensembl_gene_id"))%>%
  filter(chromosome_name%in%c(as.character(1:19),"X","Y"),gene_biotype%in%c("protein_coding"))
K913df_filtold<-K913df_old%>%filter(abs(log2FoldChange)>=lfc,padj<=pval)
```

Klf9/13 Volcano plot
```{r}
volcanooo<-function(infile,lfc=1,pval=0.05,genes=NULL){
  if(!is.null(genes)){infile<-mutate(infile,isGene=ifelse(external_gene_name%in%genes,T,F))
  } else {infile<-mutate(infile,isGene=F)}
  temp<-infile%>%filter(!is.na(padj))%>%
    mutate(Effect=case_when(log2FoldChange>=lfc & padj<pval ~ "Increase",
                             log2FoldChange<=-lfc & padj<pval ~ "Decrease",
                             TRUE ~ "No Change"))
  temptab<-table(temp$Effect)
  fig1<-temp%>%ggplot(aes(x=log2FoldChange,y=-log10(padj)))+geom_point(aes(col=Effect))+
  scale_color_manual(breaks = c("Increase","Decrease","No Change"),
                     values=alpha(c("red", "blue","gray30"),0.5),
                     labels=c(paste0("Increase (n=",temptab[2],")"),
                              paste0("Decrease (n=",temptab[1],")"),"No Change"))+
  geom_point(data=subset(temp,isGene==T),shape=21,aes(x=log2FoldChange,y=-log10(padj),fill=Effect),
             color="black",stroke=1.2)+
    geom_text_repel(data=subset(temp,isGene==T),
                    aes(x=log2FoldChange,y=-log10(padj),label=external_gene_name))+
  scale_fill_manual(breaks = c("Increase","Decrease","No Change"),
                     values=alpha(c("red", "blue","gray30"),0.5))+
  ylab(expression(-log[10](padj)))+
    xlab(expression(log[2]~FoldChange))+
  xlim(c(-10,10))+ theme_classic()+
    geom_hline(yintercept = -log10(pval),linetype="dashed")+
    geom_vline(xintercept = c(-lfc,lfc),linetype="dashed")+
  theme(axis.title=element_text(size=15),axis.text = element_text(size=15),
        legend.text = element_text(size=15),legend.title = element_text(size=18))+guides(fill="none")
  return(fig1)
}

volcanooo(K913df_old,genes=c("Klf9","Klf13"))
```

Volcano plots for the others
```{r}
KLF9<-lfcShrink(ddsk913, coef="Target_Klf9_vs_ScrR",type="apeglm")

K9df<-as.data.frame(KLF9)%>%rownames_to_column("ENSID")%>%
  left_join(bmart1,by = c("ENSID"="ensembl_gene_id"))%>%
  filter(chromosome_name%in%c(as.character(1:19),"X","Y"),gene_biotype%in%c("protein_coding"))
K9df_filt<-K9df%>%filter(abs(log2FoldChange)>=lfc,padj<=pval)

KLF13<-lfcShrink(ddsk913, coef="Target_Klf13_vs_ScrR",type="apeglm")

K13df<-as.data.frame(KLF13)%>%rownames_to_column("ENSID")%>%
  left_join(bmart1,by = c("ENSID"="ensembl_gene_id"))%>%
  filter(chromosome_name%in%c(as.character(1:19),"X","Y"),gene_biotype%in%c("protein_coding"))
K13df_filt<-K13df%>%filter(abs(log2FoldChange)>=lfc,padj<=pval)

volcanooo(K9df,gene="Klf9")
volcanooo(K13df,gene="Klf13")
```


Heatmap function
```{r}
heatmapplus<-function(countmat,target=NA,DEGset=NA,motifs=NULL,AvgExp=FALSE,LFC=FALSE,rnames=FALSE,
                      clustmeth="ward.D2",
                      hmcols=colorRampPalette(rev(brewer.pal(n=10,name="RdYlBu")))(100),
                      Candgenes=NA,motif_title="Motif Score",col=c("blue4","salmon2","firebrick2","sandybrown"),
                      col2=c("PRGn","YlGn"),returnraw=FALSE,...){
  countmat_filt<-countmat%>%rownames_to_column("ENSID")%>%filter(ENSID%in%DEGset$ENSID)%>%mutate(GeneID=make.unique(GeneID))%>%
    column_to_rownames("GeneID")
  countmat_filt<-dplyr::select(countmat_filt,starts_with(c("Scr",target)))
  scale_counts_filt<-apply(countmat_filt,1,FUN=function(x) (x-mean(x))/sd(x))
  mat<-as.data.frame(t(scale_counts_filt))
  collabs=HeatmapAnnotation(Target=str_remove(names(mat),pattern = ' [A-Z]+$'),
                            col=list(Target=c("Scramble"=col[1],"Klf9"=col[2],"Klf13"=col[4],"Klf9/13"=col[3])),
                            show_annotation_name=FALSE)
  ha = rowAnnotation(foo = anno_mark(at = match(Candgenes,rownames(mat)), 
                                     labels = Candgenes,labels_gp = gpar(fontsize = 10)))
  fig<-Heatmap(mat,col = hmcols,name="Normalized Expression",show_column_names = TRUE,
        clustering_distance_rows = "pearson",
        show_row_names = FALSE,
        clustering_method_rows = clustmeth,
        top_annotation = collabs,column_names_rot=45,cluster_columns = F,
        heatmap_legend_param = list(direction="horizontal"),...)
  if(AvgExp==TRUE){
    AvgExp<-dplyr::select(DEGset,c("external_gene_name","baseMean"))%>%
    column_to_rownames("external_gene_name")%>%mutate(baseMean=log2(baseMean))
    AvgExp<-AvgExp[match(rownames(mat),rownames(AvgExp)),]
    fig<-fig+Heatmap(AvgExp,cluster_columns =FALSE,show_row_names = FALSE,
                show_column_names = FALSE,cluster_rows=FALSE,
                col= colorRampPalette((brewer.pal(11, "PuRd")))(100),width=unit(5,"mm"),
                heatmap_legend_param = list(title="Avg. Log Expression",direction="horizontal"))} 
  if(LFC==TRUE){
    LFC<-dplyr::select(DEGset,c("external_gene_name","log2FoldChange"))%>%
    column_to_rownames("external_gene_name")
    LFC<-LFC[match(rownames(mat),rownames(LFC)),]
    fig<-fig+Heatmap(LFC,cluster_columns =FALSE,show_row_names = FALSE,
                show_column_names = FALSE,cluster_rows=FALSE,
                col=colorRamp2(c(-4, 0, 4), c("red", "white", "green")),width=unit(5,"mm"),
                heatmap_legend_param = list(title="log2FoldChange",direction="horizontal"))}
  if(!is.null(motifs)) {
      motifs<-motifs$Pospeaks[match(rownames(mat),motifs$external_gene_name)]
      motifs[is.na(motifs)]<-0
      fig<-fig+Heatmap(motifs,cluster_columns =FALSE,show_row_names = FALSE,
                show_column_names = FALSE,cluster_rows=FALSE,
                col=colorRamp2(c(0, 2.5, 5), c("white", "darkolivegreen2", "darkolivegreen4")),width=unit(5,"mm"),
                heatmap_legend_param = list(title=motif_title,direction="horizontal"))} 
  out<-fig+ha
  if(returnraw){return(fig)
    } else {
  return(out)
    }
}
```

OLD KLF Motif enrichment in DEG Promoters ((v all genes)
```{r}
shiftfunction<-function(Peak_df,FIMO_input,reduce=TRUE,motifsize=18,pfilter=FALSE){
  PeakGR<-GRanges(Peak_df)
  if(reduce){
    FIMO_input<-FIMO_input[!duplicated(FIMO_input[,3:5]),]
  }
  if(!isFALSE(pfilter)){
    FIMO_input<-FIMO_input%>%filter(q.value<=pfilter)
  }
  true_mo<-PeakGR[match(FIMO_input$sequence_name,mcols(PeakGR)$Peak_id)]
  mcols(true_mo)<-cbind(mcols(true_mo),FIMO_input[,7:10])
  true_mo<-resize(true_mo,width=FIMO_input$stop)
  true_mo<-resize(true_mo,motifsize,"end")
  names(true_mo)<-NULL
  return(true_mo)
}

#Load in FIMO output, peaks locations, and DEG set
load("~/Documents/ATACseq_Project/InternalPeaks/newpeaksRK/AllProPeaks.Rda")
bmart2<-getBM(attributes = c("external_gene_name","ensembl_gene_id"),
              mart = ensembl)
input<-AllProPeaks
FIMO_NEG<-read.table("~/Documents/MotifAnalysis/FIMO/KLF_JASPAR_root/fimo.tsv",
                     sep = "\t",header = TRUE)
Peaks_in<-input%>%mutate(Peak_id=make.unique(external_gene_name))


#Align the two (map motif postiion within peak)
Peakpos<-shiftfunction(Peak_df = Peaks_in,FIMO_input = FIMO_NEG)
Peakpos<-Peakpos[Peakpos$q.value<=0.05]


#Setup for Fisher Test(s)
motifbin<-countOverlaps(GRanges(input),Peakpos)>0
input$Motif<-motifbin
temptab<-input%>%group_by(fixedGene)%>%summarise(Pospeaks=sum(Motif),Allpeaks=n())%>%
  left_join(bmart2,by=c("fixedGene"="ensembl_gene_id"))
HmapMotifs<-temptab%>%filter(!is.na(external_gene_name))
#Significance of DEG vs Not
DEGin<-filter(K913df_filtold,log2FoldChange>=1)
Motiftable<-temptab%>%
  mutate(IsDEG=ifelse(external_gene_name%in%DEGin$external_gene_name,"DEG","Not_DEG"))%>%
  group_by(IsDEG)%>%summarize(motif=sum(Pospeaks),no_motif=sum(Allpeaks)-sum(Pospeaks))%>%
  column_to_rownames("IsDEG")
print(Motiftable)
#Test for Enrichment (less = Control<DEG)!
fisher.test(Motiftable,alternative = "greater")
```
Make GRanges for overlap testing
```{r}
mart<-useMart(host="https://jan2020.archive.ensembl.org",biomart = "ENSEMBL_MART_ENSEMBL",dataset = "mmusculus_gene_ensembl")
bmart2<-getBM(attributes = c("ensembl_gene_id","ensembl_transcript_id","transcript_start",
"transcript_end","chromosome_name","strand","transcript_biotype","transcript_gencode_basic"), mart = mart)

TSSfrombmart<-function(bmart,extend=c(100000,100000),reduce=TRUE,filter=TRUE){
  bmart<-bmart%>%filter(chromosome_name%in%c(as.character(1:19),"X","Y"))
    if(filter==TRUE){bmart<-bmart%>%filter(transcript_biotype=="protein_coding",
                                   transcript_gencode_basic=="GENCODE basic")}
  out<-GRanges(seqnames=paste0("chr",bmart$chromosome_name),
          ranges=IRanges(start=bmart$transcript_start,end=bmart$transcript_end),
          strand=ifelse(bmart$strand==1,"+","-"),
          mcols=data.frame(ENSID=bmart$ensembl_transcript_id,GeneID=bmart$ensembl_gene_id))
  outplus<-promoters(out,upstream = extend[1],downstream = extend[2])
  if(reduce==TRUE){outlist<-split(outplus,outplus$mcols.GeneID)
  outlist<-GenomicRanges::reduce(outlist)
  outplus<-unlist(outlist)
  }
  return(sort(outplus))
}

#Get genes and TSS windows as GRanges
TSSgenes<-TSSfrombmart(bmart2,extend=c(1000,200))
```


Load Subset Peak files
```{r}
load("DevATAC/P2P30_ATAC/Sacha/p2p30Peaks_rawcounts.RDa",verbose = T)
devpeakGR<-devcountGR
mcols(devpeakGR)<-NULL

#gcountsnorm<-sweep(as.matrix(mcols(devcountGR)),2,FUN="/",STATS=as.numeric(apply(mcols(devcountGR),2,sum)/1e6))
#devpeakGRfilt<-devpeakGR[apply(log2(gcountsnorm),1,max)>1]

peaksin<-devpeakGR
ovtest<-findOverlaps(TSSgenes,peaksin)
TSSpeaks<-peaksin[ovtest@to]
TSSpeaks$geneId<-names(TSSgenes)[ovtest@from]
```

NEW KLF Motif Enrichment (using p2p30 peaks): Use this one!!
```{r}
#load("~/Documents/DevATAC/p2p30PeaksSN_JASPAR2022CORE_FIMO.rda")
Peakpos<-listy$KLF5_MA0599.1

input<-Propeaks

#Setup for Fisher Test(s)
motifbin<-countOverlaps(input,Peakpos)>0
input$Motif<-motifbin
temptab<-as.data.frame(input)%>%group_by(geneId)%>%summarise(Pospeaks=sum(Motif),Allpeaks=n())%>%
  left_join(bmart1,by=c("geneId"="ensembl_gene_id"))
HmapMotifs<-temptab%>%filter(!is.na(external_gene_name))
#Significance of DEG vs Not
DEGin<-filter(K913df_filtold,log2FoldChange>=1)
Motiftable<-temptab%>%
  mutate(IsDEG=ifelse(external_gene_name%in%DEGin$external_gene_name,"DEG","Not_DEG"))%>%
  group_by(IsDEG)%>%summarize(motif=sum(Pospeaks),no_motif=sum(Allpeaks)-sum(Pospeaks))%>%
  column_to_rownames("IsDEG")
print(Motiftable)
#Test for Enrichment (less = Control<DEG)!
fisher.test(Motiftable,alternative = "greater")
```

```{r}
library(ggrepel)
volcanoooTF<-function(infile,pval=0.01,motifs=NA,allpostext=F,JASPARID=T){
  if(JASPARID){infile$TFID<-str_split(infile$TFID,pattern = "_",simplify = T)[,1]}
  if(!is.na(motifs)){infile<-mutate(infile,isMotif=ifelse(grepl(TFID,pattern=motifs,ignore.case=T),T,F))
  } else {infile<-mutate(infile,isMotif=F)}
  temp<-infile%>%filter(!is.nan(padj))%>%
                          mutate(Effect=case_when(Enrichment>1 & padj<pval ~"Enriched(Test)",
                             Enrichment<1 & padj<pval ~ "Enriched(Control)",
                             TRUE ~ "No Change"))
  temptab<-table(temp$Effect)
  fig1<-temp%>%ggplot(aes(x=Enrichment,y=-log10(padj)))+geom_point(aes(col=Effect))+
  scale_color_manual(breaks = c("Enriched(Test)","Enriched(Control)","No Change"),
                     values=ggplot2::alpha(c("red", "blue","gray30"),0.5))+
  geom_point(data=subset(temp,isMotif==T),shape=21,aes(x=Enrichment,y=-log10(padj),fill=Effect),
             color="black",stroke=1.2)+
  scale_fill_manual(breaks = c("Enriched(Test)","Enriched(Control)","No Change"),
                     values=c("red", "blue","gray30"))+
    geom_text_repel(data=subset(temp,isMotif==T),
                    aes(x=Enrichment,y=-log10(padj),label=TFID))+
  ylab(expression(-log[10](p.val)))+
    xlab("Enrichment")+
    theme_classic()+
    geom_hline(yintercept = -log10(pval),linetype="dashed")+
     geom_vline(xintercept = 1,linetype="dashed")+
  theme(axis.title=element_text(size=15),axis.text = element_text(size=15),legend.position = "bottom",
        legend.text = element_text(size=15),legend.title = element_text(size=18))+guides(fill="none")
  if(allpostext==T){
    fig1<- fig1 + geom_text_repel(data=subset(temp,Effect=="Enriched(Test)"),
                                  aes(x=Enrichment,y=-log10(padj),label=TFID))
  }
  return(fig1)
}

contingency_mat<-function(col,hitmat=temp){
  pos_test<-sum(col[hitmat])
  neg_test<-length(col[hitmat])-pos_test
  pos_cont<-sum(col[!hitmat])
  neg_cont<-length(col[!hitmat])-pos_cont
  out<-matrix(c(pos_test,neg_test,pos_cont,neg_cont),nrow=2,ncol=2,byrow=T)
  return(out)
}
```

All Motif Enrichment
```{r}
#load("~/Documents/DevATAC/p2p30PeaksSN_JASPAR2022CORE_FIMO.rda")
#load("~/Documents/DevATAC/p2p30PeaksSN_JASPAR2022center_FIMO.rda")
peaksin<-Propeaks
#peaksin<-TSSpeaks[TSSpeaks$geneID%in%bfilt$ensembl_gene_id[bfilt$external_gene_name%in%TPMfilt]]

testout<-sapply(listy,function(x) countOverlaps(peaksin,x))

listx<-list()
for(i in seq(ncol(testout))){
  listx[[i]]<-testout[,i]>0
}

DEGin<-filter(K913df_filtold,log2FoldChange>=1)
temp<-peaksin$geneId%in%DEGin$ENSID

contabs<-lapply(listx,FUN=contingency_mat)
names(contabs)<-names(listy)

Chitest_pos<-data.frame(Enrichment=sapply(contabs,FUN=function(x) (x[1,1]/sum(x[1,]))/(x[2,1]/sum(x[2,]))))%>%
  mutate(TFID=names(contabs),Pval=sapply(contabs,FUN=function(x) chisq.test(x)$p.value))%>%
  mutate(padj=p.adjust(Pval,"BH"))

Fishtest<-lapply(contabs,FUN=function(x) fisher.test(x))
Fishtest_pos<-data.frame(Enrichment=sapply(Fishtest,function(x) x$estimate),
                         Pval=sapply(Fishtest,function(x) x$p.value))%>%
                           mutate(TFID=names(contabs),padj=p.adjust(Pval,method="BH"))

volcanoooTF(Chitest_pos,motifs="Klf")
```


Plot heatmap, split w/ labels
```{r}
library(ComplexHeatmap)
library(RColorBrewer)
library(circlize)

IonChannels<-paste(c("^Gri[akn]+[1-9]*","^Scn[1-9]*[ab]+","^Kcn[mnujtkabcdfghqsv]+[1-9]*","^Cacn[agb]+[1-9]*","^Gabr[abdegpqr]+","^Hcn[1-4]*"),collapse = "|")
allTFs<-read.table("~/Documents/Misc/Mus_musculus_TF.txt",header = TRUE)
axongenes<-c("Dpysl5","L1cam","Gap43","Rac3","Tubb2b","Tubb3","Hcn3","Cacng4","Ablim3","Dpysl3","Ndrg1","Itgb4","Atat1","Dpysl2","Mical1","Hook2","Tppp3")
KLFbestgenes<-c("S100a10","Dpysl3","Dpysl5","Tubb2b","Gap43","Tppp3","Rac3","Bcam","Ppp1r18","Scn9a","Scn3a","Scn3b","Hcn3","Atat1","Hook2","Maged2","L1cam","N4bp3","Plppr1","Prmt2","Trpm3","Pcdhgc4","Zcchc12","Ablim3","Ddah2","Cacng4","Tubb3","Tubb5","Tubb1a","Stmn2","Eps8","Coro2a","Myo1c","Iqgap1","Mturn","Rhov","Pik3dc","Kcnip3","Cobl","Nectin2","Plxna3","Diaph3","Syndig1l","Plppr3","Klf9","Klf13","Klf6","Klf7")

#normalized counts
countmat_rlg<-as.data.frame(assay(rlg_batch))%>%
  mutate(GeneID=bmart1$external_gene_name[match(rownames(.),bmart1$ensembl_gene_id)])
#names(countmat_rlg)<-c("Klf13 A","Klf13 B","Klf13 C","Klf9 A","Klf9 B","Klf9 C","Klf9/13 A", "Klf9/13 B","Klf9/13 C","Scramble A","Scramble B","Scramble C","Scramble D","GeneID")
names(countmat_rlg)<-c("Klf13 A","Klf13 B","Klf13 C","Klf13 D","Klf9 A","Klf9 B","Klf9 C","Klf9 D","Klf9/13 A", "Klf9/13 B","Klf9/13 C","Klf9/13 D","Scramble A","Scramble B","Scramble C","Scramble D","GeneID")

DEGset<-K913df_filtold
Candgenes<-c(intersect(DEGset$external_gene_name,c("Klf9","Klf13",axongenes)),
                grep(DEGset$external_gene_name,pattern=IonChannels,value=TRUE))

y<-heatmapplus(countmat_rlg,target=c("Klf13","Klf9","Klf9/13"),DEGset = DEGset,Candgenes=Candgenes,motifs=temptab,motif_title = "KLF Motif Counts",row_km=2,show_parent_dend_line=F,row_title=c("",""))
#save as pdf (avoid mislabeling rows)
#pdf(file="file/path/image.pdf",width=8,height=6)
draw(y,heatmap_legend_side="bottom")
#dev.off()
```

Normalized Trend Line Plots
```{r}
DEGset<-K913df_filtold%>%filter(log2FoldChange>=1)

countmat_filt<-countmat_rlg%>%filter(GeneID%in%DEGset$external_gene_name)
scale_counts_filt<-apply(countmat_filt[,-17],1,FUN=function(x) (x-mean(x))/sd(x))
mat<-as.data.frame(t(scale_counts_filt))%>%rownames_to_column("ENSID")%>%pivot_longer(-ENSID,names_to = "Sample")%>%
  mutate(Celltype=str_remove(Sample," [A-Z]$"))%>%group_by(ENSID,Celltype)%>%summarize(avgexp=mean(value))%>%
  mutate(Celltype=factor(Celltype,levels = c("Scramble","Klf13","Klf9","Klf9/13")))%>%
  group_by(Celltype)
plotup<-mat%>%ggplot(aes(x=Celltype,y=avgexp))+geom_line(aes(group=ENSID),alpha=0.1,size=1)+  stat_summary(fun=mean, group=1, geom="line", colour="blue",size=2)+stat_summary(fun=mean,fun.min=function(x) mean(x)-sd(x),fun.max=function(x) mean(x)+sd(x),geom="errorbar", colour="blue",width=0.2,size=2)+labs(x="gRNA Target", y="Normalzied Expresion")+theme_bw()+theme(axis.title =element_text(size=17),axis.text =element_text(size=15))


DEGset<-K913df_filtold%>%filter(log2FoldChange<=-1)

countmat_filt<-countmat_rlg%>%filter(GeneID%in%DEGset$external_gene_name)
scale_counts_filt<-apply(countmat_filt[,-17],1,FUN=function(x) (x-mean(x))/sd(x))
mat<-as.data.frame(t(scale_counts_filt))%>%rownames_to_column("ENSID")%>%pivot_longer(-ENSID,names_to = "Sample")%>%
  mutate(Celltype=str_remove(Sample," [A-Z]$"))%>%group_by(ENSID,Celltype)%>%summarize(avgexp=mean(value))%>%
  mutate(Celltype=factor(Celltype,levels = c("Scramble","Klf13","Klf9","Klf9/13")))%>%
  group_by(Celltype)
plotdown<-mat%>%ggplot(aes(x=Celltype,y=avgexp))+geom_line(aes(group=ENSID),alpha=0.1,size=1)+  stat_summary(fun=mean, group=1, geom="line", colour="red",size=2)+stat_summary(fun=mean,fun.min=function(x) mean(x)-sd(x),fun.max=function(x) mean(x)+sd(x),geom="errorbar", colour="red",width=0.2,size=2)+labs(x="gRNA Target", y="Normalzied Expresion")+theme_bw()+theme(axis.title =element_text(size=18),axis.text =element_text(size=16))

```


```{r}
quickplotnew<-function(TPMmatrix,celltypes=c("Klf9","Klf13","Klf913"),genes=NA,logscale=F){
  mat<-TPMmatrix%>%filter(gene_id%in%genes)%>%
    dplyr::select(contains("Scr"),contains(celltypes),"gene_id")%>%
    pivot_longer(cols = -gene_id)%>%mutate(Target=str_remove(name,"(GFP)?_?[A-Z][0-9]?$"))
  mat$Target[mat$Target=="ScrR"]="Scramble gRNA"
  if(logscale){mat<-mat%>%mutate(value=log2(value)+1)}
  out<-mat%>%group_by(gene_id,Target)%>%
    mutate(mean=mean(value),std=sd(value),
           Target=factor(Target,levels = c("Scramble gRNA","Klf13","Klf9","Klf913")))%>%
    ggplot(aes(x=gene_id,y=mean,fill=Target))+
    geom_bar(stat="summary",position="dodge",color="black")+
    geom_errorbar(aes(ymin=mean-std,ymax=mean+std),position=position_dodge(.9),width=0.2)+
    geom_point(aes(y=value),position=position_dodge(width=0.9),show.legend = F)+theme_classic()+
    scale_fill_manual(values=c("blue4","sandybrown","salmon2","firebrick2"))+labs(y="Normalized Counts (TPM)",x="Gene",fill="Target")+
    scale_y_continuous(expand=c(0,NA))
  if(logscale){out<-out+labs(y="Normalized Counts (Log2TPM)")}
  return(out)
}

TPMmat<-read.csv("~/Documents/Counts/KLF_TPM.csv",row.names = 2)%>%dplyr::select(-X.1)
TPMmat<-TPMmat[,-c(4,9,16)]

topgenes<-c("Ablim3","Atat1","Cacng4","Dkkl1","Dpysl3","Maged2","Tubb2b","Zcchc12","Hook2","Tppp3")
targetplot<-quickplotnew(TPMmatrix = TPMmat,celltypes = c("Klf9","Klf13"),genes = topgenes)+scale_fill_manual(values=c("blue4","sandybrown","salmon2","firebrick2"))+labs(y="Normalized Counts (TPM)",x="Gene",fill="Target")+theme(legend.position = "right",axis.title =element_text(size=15),axis.text.x =element_text(size=15,angle = 45,vjust = 0.6))
```

Venn Diagram plots
```{r}
library(eulerr)
K913_13_all<-length(intersect(K913df_filtold$ENSID,K13df_filt$ENSID))
K913_13_up<-length(intersect(K913df_filtold[K913df_filtold$log2FoldChange>=1,"ENSID"],K13df_filt[K13df_filt$log2FoldChange>=1,"ENSID"]))
K913_13_down<-length(intersect(K913df_filtold[K913df_filtold$log2FoldChange<=-1,"ENSID"],K13df_filt[K13df_filt$log2FoldChange<=-1,"ENSID"]))
K9_13_all<-length(intersect(K13df_filt$ENSID,K9df_filt$ENSID))
K9_13_up<-length(intersect(K9df_filt[K9df_filt$log2FoldChange>=1,"ENSID"],K13df_filt[K13df_filt$log2FoldChange>=1,"ENSID"]))
K9_13_down<-length(intersect(K9df_filt[K9df_filt$log2FoldChange<=-1,"ENSID"],K13df_filt[K13df_filt$log2FoldChange<=-1,"ENSID"]))
K913_9_all<-length(intersect(K913df_filtold$ENSID,K9df_filt$ENSID))
K913_9_up<-length(intersect(K913df_filtold[K913df_filtold$log2FoldChange>=1,"ENSID"],K9df_filt[K9df_filt$log2FoldChange>=1,"ENSID"]))
K913_9_down<-length(intersect(K913df_filtold[K913df_filtold$log2FoldChange<=-1,"ENSID"],K9df_filt[K9df_filt$log2FoldChange<=-1,"ENSID"]))

klfcomp<-data.frame(Klf9=c(length(K9df_filt$ENSID),length(K9df_filt[K9df_filt$log2FoldChange>=1,"ENSID"]),length(K9df_filt[K9df_filt$log2FoldChange<=-1,"ENSID"]),
                           K9_13_all,K9_13_up,K9_13_down,
                           K913_9_all,K913_9_up,K913_9_down),
                    Klf13=c(K9_13_all,K9_13_up,K9_13_down,
                            length(K13df_filt$ENSID),length(K13df_filt[K13df_filt$log2FoldChange>=1,"ENSID"]),length(K13df_filt[K13df_filt$log2FoldChange<=-1,"ENSID"]),
                            K913_13_all,K913_13_up,K913_13_down),
                    Klf913=c(K913_9_all,K913_9_up,K913_9_down,
                             K913_13_all,K913_13_up,K913_13_down,
                             length(K913df_filtold$ENSID),length(K913df_filtold[K913df_filtold$log2FoldChange>=1,"ENSID"]),length(K913df_filtold[K913df_filtold$log2FoldChange<=-1,"ENSID"])))

rownames(klfcomp)<-c("Klf9_all","Klf9_up","Klf9_down","Klf13_all","Klf13_up","Klf13_down","Klf913_all","Klf913_up","Klf913_down")


vdall<-euler(c("Klf9"=4,"Klf13"=4,"Klf9/13"=187,"Klf9&Klf9/13"=24,"Klf9&Klf13"=0,"Klf13&Klf9/13"=1,"Klf9&Klf13&Klf9/13"=0))
vdup<-euler(c("Klf9"=1,"Klf13"=3,"Klf9/13"=160,"Klf9&Klf9/13"=21,"Klf9&Klf13"=0,"Klf13&Klf9/13"=0,"Klf9&Klf13&Klf9/13"=0))
vddown<-euler(c("Klf9"=3,"Klf13"=3,"Klf9/13"=27,"Klf9&Klf9/13"=3,"Klf9&Klf13"=0,"Klf13&Klf9/13"=1,"Klf9&Klf13&Klf9/13"=0))

plot(vd,counts=TRUE,font=7,labels=NULL,quantities=list(cex=2),cex=1,alpha=0.5,fill=c("turquoise3","green3","firebrick2"))
plot(vdup,counts=TRUE,font=7,labels=NULL,quantities=list(cex=2.5),cex=1,alpha=0.5,fill=c("turquoise3","green3","firebrick2"))
plot(vddown,counts=TRUE,font=7,labels=NULL,quantities=list(cex=3),cex=1,alpha=0.5,fill=c("turquoise3","green3","firebrick2"))
```

```{r}
countmat<-read.csv("~/Documents/DevATAC/P2P30_RNA/p2p30_mm10_counts.csv")%>%mutate(X=make.unique(X))%>%column_to_rownames("X")
countmat<-countmat[,c(1:4,13:16,21:28)]
#countmat<-countmat%>%dplyr::select(contains("L4_RORht"))
Age<-ifelse(grepl(names(countmat),pattern="p2"),"P2","P30")
Layer<-ifelse(grepl(names(countmat),pattern="L4"),"L4","L6")
sampledf<-data.frame(Age=Age,Layer=Layer)
rownames(sampledf)<-colnames(countmat)

#countmat<-read.csv("~/Documents/DevATAC/P2P30_RNA/p2p30_mm10_counts_s1.csv")%>%mutate(X=make.unique(X))%>%column_to_rownames("X")
#countmat<-countmat%>%dplyr::select(contains("L4_RORht"))%>%dplyr::select(matches("p2|p30|p7"))
#Age<-ifelse(grepl(names(countmat),pattern="p2"),"P2","P30")
#Layer<-ifelse(grepl(names(countmat),pattern="L4"),"L4","L6")
#sampledf<-data.frame(Age=Age,Layer=Layer)
#rownames(sampledf)<-colnames(countmat)

DESeq <- DESeqDataSetFromMatrix(countmat,
                                   colData = sampledf,
                                   design = ~ Layer+Age)
DESeq$Target<-relevel(DESeq$Age,"P2")
keep <- rowSums(counts(DESeq)) >= 10
filt <- DESeq[keep,]
Devdds <- DESeq(filt)

DevLFC<-lfcShrink(Devdds, coef="Age_P30_vs_P2",type="apeglm")

Devdf<-as.data.frame(DevLFC)%>%rownames_to_column("ENSID")%>%
  left_join(bmart1,by = c("ENSID"="external_gene_name"))

Devrlg<-as.data.frame(assay(rlog(Devdds,blind = FALSE)))
```
Ephys - mini plot
```{r}

```

Ephys - FI plot
```{r}

```


line plots of all Klf9/13 targets across development (p2 > p7 > p30)
```{r}
bmart3<-getBM(attributes = c("ensembl_gene_id_version","external_gene_name"), mart = ensembl)
DevTPM<-read.csv("~/Documents/Counts/p2p30/p2p30_L4_RefGenes.csv",row.names = 1)
klfensnames<-c("Klf13","Klf9"))
DEGset<-K913df_filtold%>%filter(log2FoldChange>=1)

countmat_filt<-DevTPM%>%filter(external_gene_name%in%c(DEGset$external_gene_name,"Klf9","Klf13"))
scale_counts_filt<-apply(countmat_filt[,-13],1,FUN=function(x) (x-mean(x))/sd(x))
mat<-as.data.frame(t(scale_counts_filt))%>%rownames_to_column("ENSID")%>%pivot_longer(-ENSID,names_to = "Sample")%>%
  mutate(Age=str_extract(Sample,pattern="p[0-9]+"))%>%mutate(Age=str_replace(Age,pattern="p",replacement="P"))%>%
  mutate(Age=fct_relevel(Age,"P2","P7","P30"))%>%group_by(ENSID,Age)%>%summarize(avgexp=mean(value))%>%
  group_by(Age)
ploty<-mat%>%filter(!ENSID%in%klfensnames$ensembl_gene_id_version)%>%ggplot(aes(x=Age,y=avgexp))+geom_line(aes(group=ENSID),alpha=0.1)+  stat_summary(fun=mean, group=1, geom="line", colour="blue",linewidth=1)+stat_summary(fun=mean,fun.min=function(x) mean(x)-sd(x),fun.max=function(x) mean(x)+sd(x),geom="errorbar", colour="blue",width=0.2,linewidth=1)+labs(x="Age", y="Normalzied Expresion")+theme_bw()+theme(axis.title =element_text(size=17),axis.text =element_text(size=15))
ploty+geom_line(data=subset(mat,ENSID%in%klfensnames),aes(x=Age,y=avgexp,col=ENSID,group=ENSID),linewidth=1.2,linetype="dashed")+scale_color_manual(name="Gene",values=c("red3","tomato"),labels=c("Klf9","Klf13"))+theme(legend.position = "bottom")
```

Dot plot of P2 v P30, K913 v scramble
```{r}
Dev913<-K913df_old[,c(1,3,5,6,7)]%>%inner_join(Devdf[,c(1,3,6)],by=c("external_gene_name"="ENSID"),suffix=c("_KLF","_Dev"))%>%
  mutate(DEG=ifelse(ENSID%in%K913df_filtold$ENSID,"Klf9/13 Target","Not Target"))%>%dplyr::filter(!external_gene_name%in%c("Klf9","Klf13"))

out<-ggplot(Dev913,aes(x=log2FoldChange_Dev,y=log2FoldChange_KLF))+geom_point(aes(col=DEG,alpha=DEG)) + scale_color_manual(values = c("Not Target" = "black", "Klf9/13 Target" = "firebrick2")) + 
  labs(x="log2FoldChange(P2 v P30)",y="log2FoldChange(Klf9/13 v Scramble)",col="")+geom_vline(xintercept = 0,linetype="dashed")+geom_hline(yintercept = 1,linetype="dashed")+
  geom_hline(yintercept = -1,linetype="dashed") + theme_bw() + geom_smooth(method=lm,se = T,col="grey40")+stat_cor(aes(label=..r.label..),label.x=4,label.y=4,size=5)+
  stat_smooth(data=subset(Dev913,DEG=="Klf9/13 Target"),col = "firebrick2",method=lm,fullrange = T) + 
  stat_cor(data=subset(Dev913,DEG=="Klf9/13 Target"),col="firebrick2",label.x=4,label.y=3.5,size=5)+scale_alpha_manual(values=c(1,0.4))+
  theme(axis.title =element_text(size=16),axis.text =element_text(size=13))

```

Klf67 Age qPCR
```{r}
chk<-fread("~/Documents/qPCR/Klf67913_Age/K67Ageqpcr.csv")
KLF<-chk%>%dplyr::select(SampID,FC_K6_P10,FC_K7_P10,Virus,Age)%>%
  pivot_longer(cols=c(FC_K6_P10,FC_K7_P10),names_to = "Target",values_to = "FC")%>%
  mutate(Target=ifelse(Target=="FC_K6_P10","Klf6","Klf7"))%>%arrange(Target,desc(Virus),Age)%>%
  group_by(Virus,Age,Target)%>%mutate(fcmean=mean(FC))
n<-length(unique(KLF$SampID))
KLF$controlmean<-c(rep(KLF$fcmean[1],n),rep(KLF$fcmean[n+1],n))
KLF<-KLF%>%mutate(fcnorm=FC/controlmean)%>%mutate(normmean=mean(fcnorm),fcsd=sd(fcnorm))%>%mutate(Group=paste(Virus, Age, sep=" "))%>%
  mutate(Group=factor(Group,levels = c("Scr-GFP P10-12","Klf67-GFP P10-12","Scr-GFP P18-20","Klf67-GFP P18-20")))
y<-ggplot(KLF,aes(x=Target,y=normmean,fill=Group))+geom_bar(stat="summary",position="dodge",color="black")+
  geom_point(aes(x=Target,y=FC),position = position_dodge(width=0.9),show.legend = F)+
  geom_errorbar(aes(ymin=normmean-fcsd,ymax=normmean+fcsd,x=Target),position = position_dodge(width=0.9),width=0.2)+
  scale_y_continuous(expand=c(0,0),limit=c(0,1.75))+theme_classic()+scale_fill_brewer(palette = "Paired")+
  labs(x="Measured Gene",y="Relative mRNA Levels",fill="gRNA Target")+theme(legend.position = "right",text=element_text(size=16))+
  geom_hline(yintercept = 1,linetype="dashed")+theme(text=element_text(size=18),legend.position = "right",legend.text = element_text(size=12))

```

Klf67 RNA-seq by Age
```{r}
#load("~/Documents/CRISPRi/SeqAnalysis/KlfP10P20_DESeq2.RDa")
lfc<-log2(2)
pval<-0.05
KLF67_20<-lfcShrink(dds, contrast = c("Group","Klf67_P20","ScrGFP_P20"),type="ashr")

K67df_20<-as.data.frame(KLF67_20)%>%rownames_to_column("ENSID")%>%
  left_join(bmart1,by = c("ENSID"="ensembl_gene_id"))%>%
  filter(chromosome_name%in%c(as.character(1:19),"X","Y"),gene_biotype%in%c("protein_coding"))
K67df20_filt<-K67df_20%>%filter(abs(log2FoldChange)>=lfc,padj<=pval)

KLF67_10<-lfcShrink(dds, contrast = c("Group","Klf67_P10","ScrGFP_P10"),type="ashr")

K67df_10<-as.data.frame(KLF67_10)%>%rownames_to_column("ENSID")%>%
  left_join(bmart1,by = c("ENSID"="ensembl_gene_id"))%>%
  filter(chromosome_name%in%c(as.character(1:19),"X","Y"),gene_biotype%in%c("protein_coding"))
K67df10_filt<-K67df_10%>%filter(abs(log2FoldChange)>=lfc,padj<=pval)

K67p10_vol<-volcanooo(K67df_10,genes=c("Klf6","Klf7"))
K67p20_vol<-volcanooo(K67df_20,genes=c("Klf6","Klf7"))
```

```{r}
heatmapplus2<-function(countmat,target=NA,DEGset=NA,motifs=NULL,AvgExp=FALSE,LFC=FALSE,rnames=FALSE,
                      clustmeth="ward.D2",remix=TRUE,
                      hmcols=colorRampPalette(rev(brewer.pal(n=10,name="RdYlBu")))(100),
                      Candgenes=NA,motif_title="Motif Score",col=c("lightblue2","blue4","firebrick2","lightgreen","forestgreen"),
                      col2=c("PRGn","YlGn"),returnraw=FALSE,...){
  countmat_filt<-countmat%>%rownames_to_column("ENSID")%>%filter(ENSID%in%DEGset$ENSID)%>%
    column_to_rownames("GeneID")%>%dplyr::select(-ENSID)
  if(remix==TRUE){countmat_filt<-dplyr::select(countmat_filt,starts_with(c("Scr",target)))}
  scale_counts_filt<-apply(countmat_filt,1,FUN=function(x) (x-mean(x))/sd(x))
  mat<-as.data.frame(t(scale_counts_filt))
  collabs=HeatmapAnnotation(Target=str_remove(names(mat),pattern = ' [A-Z]+$'),
                            col=list(Target=c("Scramble P10"=col[1],"Scramble P20"=col[2],"Klf9/13 P20"=col[3],"Klf6/7 P10"=col[4],"Klf6/7 P20" = col[5])),
                            show_annotation_name=FALSE)
  ha = rowAnnotation(foo = anno_mark(at = match(Candgenes,rownames(mat)), 
                                     labels = Candgenes,labels_gp = gpar(fontsize = 10)))
  fig<-Heatmap(mat,col = hmcols,name="Normalized Expression",show_column_names = TRUE,
        clustering_distance_rows = "pearson",
        clustering_method_rows = clustmeth,
        top_annotation = collabs,column_names_rot=45,cluster_columns = F,
        heatmap_legend_param = list(direction="horizontal"),...)
  if(AvgExp==TRUE){
    AvgExp<-dplyr::select(DEGset,c("external_gene_name","baseMean"))%>%
    column_to_rownames("external_gene_name")%>%mutate(baseMean=log2(baseMean))
    AvgExp<-AvgExp[match(rownames(mat),rownames(AvgExp)),]
    fig<-fig+Heatmap(AvgExp,cluster_columns =FALSE,show_row_names = FALSE,
                show_column_names = FALSE,cluster_rows=FALSE,
                col= colorRampPalette((brewer.pal(11, "PuRd")))(100),width=unit(5,"mm"),
                heatmap_legend_param = list(title="Avg. Log Expression",direction="horizontal"))} 
  if(LFC==TRUE){
    LFC<-dplyr::select(DEGset,c("external_gene_name","log2FoldChange"))%>%
    column_to_rownames("external_gene_name")
    LFC<-LFC[match(rownames(mat),rownames(LFC)),]
    fig<-fig+Heatmap(LFC,cluster_columns =FALSE,show_row_names = FALSE,
                show_column_names = FALSE,cluster_rows=FALSE,
                col=colorRamp2(c(-4, 0, 4), c("red", "white", "green")),width=unit(5,"mm"),
                heatmap_legend_param = list(title="log2FoldChange",direction="horizontal"))}
  if(!is.null(motifs)) {
      motifs<-motifs$Pospeaks[match(rownames(mat),motifs$external_gene_name)]
      motifs[is.na(motifs)]<-0
      fig<-fig+Heatmap(motifs,cluster_columns =FALSE,
                show_column_names = FALSE,cluster_rows=FALSE,
                col=colorRamp2(c(0, 2.5, 5), c("white", "darkolivegreen2", "darkolivegreen4")),width=unit(5,"mm"),
                heatmap_legend_param = list(title=motif_title,direction="horizontal"))} 
  out<-fig+ha
  if(returnraw){return(fig)
    } else {
  return(out)
    }
}
```

```{r}
countmat_rlg<-as.data.frame(assay(rlog(dds,blind = FALSE)))%>%
  mutate(GeneID=bmart1$external_gene_name[match(rownames(.),bmart1$ensembl_gene_id)])
#names(countmat_rlg)<-c("Klf13 A","Klf13 B","Klf13 C","Klf9 A","Klf9 B","Klf9 C","Klf9/13 A", "Klf9/13 B","Klf9/13 C","Scramble A","Scramble B","Scramble C","Scramble D","GeneID")
names(countmat_rlg)<-c("Klf6/7 P10 A","Klf6/7 P10 B","Klf6/7 P10 C","Klf6/7 P10 D","Klf6/7 P20 A","Klf6/7 P20 B","Klf6/7 P20 C","Klf6/7 P20 D","Klf9/13 P20 A","Klf9/13 P20 B","Klf9/13 P20 C","Klf9/13 P20 D", "Scramble P10 A","Scramble P10 B","Scramble P10 C","Scramble P10 D","Scramble P20 A","Scramble P20 B","Scramble P20 C","Scramble P20 D","GeneID")

DEGset<-rbind(K67df10_filt)
Candgenes<-c(intersect(DEGset$external_gene_name,c(axongenes,"Klf6","Klf7")),
                grep(DEGset$external_gene_name,pattern=IonChannels,value=TRUE))
set.seed(777)
x<-heatmapplus2(countmat_rlg,target=c("Klf6/7"),DEGset = DEGset,Candgenes = Candgenes,row_km=3,row_km_repeats=100,show_parent_dend_line=F,row_title=c("","",""))
#save as pdf (avoid mislabeling rows)
#pdf(file="~/Documents/CRISPRi/Figures/K67hm_split2.pdf",width=8,height=6)
x2<-draw(x,heatmap_legend_side = "bottom")
#dev.off()

klist<-row_order(x2)
countmat_filt<-countmat_rlg%>%rownames_to_column("ENSID")%>%filter(ENSID%in%DEGset$ENSID)
temp<-data.frame(RowID=do.call(c,klist),Cluster=rep(names(klist),sapply(klist,length)))%>%arrange(RowID)
temp$ENSID<-countmat_filt$ENSID
temp$GeneID<-countmat_filt$GeneID
#Devclust<-temp%>%filter(Cluster==3)

#save as pdf (avoid mislabeling rows)

pdf(file="~/Documents/CRISPRi/Figures/K67hm_splitnew.pdf",width=8,height=6)
x<-heatmapplus2(countmat_rlg,target=c("Klf6/7"),DEGset = DEGset,Candgenes = Candgenes,motifs = temptab,
                motif_title = "KLF Motif Counts",show_parent_dend_line=F,row_title=c("","",""),row_split = factor(as.numeric(temp$Cluster),levels=c(2,3,1)))
x<-heatmapplus2(countmat_rlg,target=c("Klf6/7"),DEGset = DEGset,Candgenes = Candgenes,show_parent_dend_line=F,row_title=c("","",""),
                row_split =factor(as.numeric(temp$Cluster),levels=c(2,3,1)))
x2<-draw(x,heatmap_legend_side = "bottom")
dev.off()
```

```{r}
#Significance of DEG vs Not
DEGin<-filter(K67df10_filt,log2FoldChange<=-1)
Motiftable<-temptab%>%
  mutate(IsDEG=ifelse(external_gene_name%in%DEGin$external_gene_name,"DEG","Not_DEG"))%>%
  group_by(IsDEG)%>%summarize(motif=sum(Pospeaks),no_motif=sum(Allpeaks)-sum(Pospeaks))%>%
  column_to_rownames("IsDEG")
print(Motiftable)
#Test for Enrichment (less = Control<DEG)!
fisher.test(Motiftable,alternative = "greater")
```

Klf6/7 GO analysis
```{r}
library(clusterProfiler)
library(org.Mm.eg.db)

geneuniv<-K67df_10$ENSID
downgenes<-K67df10_filt%>%filter(log2FoldChange<=0,!external_gene_name%in%c("Klf6","Klf7"))

#downgenes<-temp%>%filter(Cluster==3,!GeneID%in%c("Klf6","Klf7"))

#upgenes<-cpcfilt%>%filter(log2FoldChange>=log2(fc),padj<=pval)
GOdown<- enrichGO(gene = downgenes$ENSID, universe = geneuniv,
                OrgDb = org.Mm.eg.db,
                ont = "ALL",minGSSize = 25,
                pAdjustMethod = "BH",keyType = "ENSEMBL",
                pvalueCutoff  = 0.05,qvalueCutoff = 0.05,
                readable = TRUE)


go_gene_matrix <- simplify(GOdown,0.5)@result %>%
  dplyr::select(Description, qvalue, geneID) %>%
  separate_rows(geneID, sep = "/") %>%      # one row per (GO term, gene) pair
  mutate(present = 1) %>%
  pivot_wider(names_from = geneID,values_from=present,values_fill = 0)

barplot<-go_gene_matrix[c(13,14,8,2,9,11,10,12,3,5,4),]%>%
  mutate(Description=factor(Description,levels=.$Description[order(.$qvalue,decreasing = T)]))%>%
  ggplot(aes(x=Description,y=-log10(qvalue)))+geom_col(color="black",linewidth = 0.25,width = 0.75,fill="lightgreen")+
  coord_flip()+scale_x_discrete(position = "top")+scale_y_continuous(expand = expansion(mult = c(0,0.1)))+theme_classic()+
  theme(axis.text.y.right = element_text(hjust = 0,face = "bold"),axis.text = element_text(size = 12),
        title = element_text(size=14,face="bold"),axis.ticks.length = unit(0.3,"cm"))

#Get top expressed genes from interesting terms for new heatmap
#Actin-related genes ( top 15)
actgenes<-unique(c(names(go_gene_matrix)[which(go_gene_matrix[10,]==1)],names(go_gene_matrix)[which(go_gene_matrix[4,]==1)]))
actgenes_top<-K67df_10%>%filter(external_gene_name%in%actgenes)%>%arrange(desc(baseMean))%>%head(15)%>%pull(external_gene_name)
#Calcium dependent genes (top 10)
cagenes<-unique(c(names(go_gene_matrix)[which(go_gene_matrix[13,]==1)],names(go_gene_matrix)[which(go_gene_matrix[14,]==1)]))
cagenes_top<-K67df_10%>%filter(external_gene_name%in%cagenes)%>%arrange(desc(baseMean))%>%head(10)%>%pull(external_gene_name)
#Axon genes (top 15)
axgenes<-names(go_gene_matrix)[which(go_gene_matrix[9,]==1)]
axgenes_top<-K67df_10%>%filter(external_gene_name%in%axgenes)%>%arrange(desc(baseMean))%>%head(15)%>%pull(external_gene_name)
Candgenes_top<-unique(c(axgenes_top,actgenes_top,cagenes_top))
```
Klf9/13 6/7 development PCA
```{r}
rlg<-rlog(dds,blind = FALSE)
tst<-plotPCA(rlg,intgroup=c("Target","Age"),ntop=500)+theme_bw()+labs(color= "Group")+
  scale_color_manual(values=c("lightgreen","forestgreen","firebrick2","lightblue2","blue4"),name="gRNA Target",
                     labels=c("Klf6/7 P10","Klf6/7 P20","Klf9/13","Scramble P10","Scramble P20"))+
  theme(axis.title =element_text(size=14),legend.position = "right")+
  geom_point(color="black",pch=21,size=4)+geom_point(aes(fill=group),size=3.5,show.legend = F)
```

Klf6/7 Development plot
```{r}
library(ggpubr)
P10P20<-lfcShrink(dds, contrast = c("Group","ScrGFP_P20","ScrGFP_P10"),type="ashr")

P10P20df<-as.data.frame(P10P20)%>%rownames_to_column("ENSID")%>%
  left_join(bmart1,by = c("ENSID"="ensembl_gene_id"))%>%
  filter(chromosome_name%in%c(as.character(1:19),"X","Y"),gene_biotype%in%c("protein_coding"))

Dev67<-K67df_10[,c(1,3,5,6,7)]%>%inner_join(Devdf[,c(1,3,6)],by=c("external_gene_name"="ENSID"),suffix=c("_KLF","_Dev"))%>%
  mutate(DEG=ifelse(ENSID%in%K67df10_filt$ENSID,"Klf6/7 Target","Not Target"))%>%dplyr::filter(!external_gene_name%in%c("Klf6","Klf7","Tmem150b"))

out<-ggplot(Dev67,aes(x=log2FoldChange_Dev,y=log2FoldChange_KLF))+geom_point(aes(col=DEG,alpha=DEG)) + scale_color_manual(values = c("Not Target" = "black", "Klf6/7 Target" = "forestgreen")) + 
  labs(x="log2FoldChange(P2 v P30)",y="log2FoldChange(Klf6/7 v Scramble, P10)",col="")+geom_vline(xintercept = 0,linetype="dashed")+geom_hline(yintercept = 1,linetype="dashed")+
  geom_hline(yintercept = -1,linetype="dashed") + theme_bw() + geom_smooth(method=lm,se = T,col="grey40")+stat_cor(aes(label=..r.label..),label.x=2.5,label.y=3.5,size=5)+
  stat_smooth(data=subset(Dev67,DEG=="Klf6/7 Target"),col = "forestgreen",method=lm,fullrange = T) + 
  stat_cor(data=subset(Dev67,DEG=="Klf6/7 Target"),col="forestgreen",label.x=2.5,label.y=3,size=5)+
  theme(axis.title =element_text(size=16),axis.text =element_text(size=13))+
  scale_alpha_manual(values=c(1,0.4))
```


```{r}
DEGset<-K67df10_filt%>%filter(log2FoldChange<=-1)
klfensnames<-c("Klf6","Klf7")

countmat_filt<-DevTPM%>%filter(external_gene_name%in%c(DEGset$external_gene_name,"Klf6","Klf7"))
scale_counts_filt<-apply(countmat_filt[,-13],1,FUN=function(x) (x-mean(x))/sd(x))
mat<-as.data.frame(t(scale_counts_filt))%>%rownames_to_column("ENSID")%>%pivot_longer(-ENSID,names_to = "Sample")%>%
  mutate(Age=str_extract(Sample,pattern="p[0-9]+"))%>%mutate(Age=str_replace(Age,pattern="p",replacement="P"))%>%
  mutate(Age=fct_relevel(Age,"P2","P7","P30"))%>%group_by(ENSID,Age)%>%summarize(avgexp=mean(value))%>%
  group_by(Age)
ploty<-mat%>%ggplot(aes(x=Age,y=avgexp))+geom_line(aes(group=ENSID),alpha=0.1)+  stat_summary(fun=mean, group=1, geom="line", colour="blue",linewidth=1)+stat_summary(fun=mean,fun.min=function(x) mean(x)-sd(x),fun.max=function(x) mean(x)+sd(x),geom="errorbar", colour="blue",width=0.2,linewidth=1)+labs(x="Age", y="Normalzied Expresion")+theme_bw()+theme(axis.title =element_text(size=17),axis.text =element_text(size=15))
ploty+geom_line(data=subset(mat,ENSID%in%klfensnames),aes(x=Age,y=avgexp,col=ENSID,group=ENSID),linewidth=1.2,linetype="dashed")+scale_color_manual(name="Gene",values=c("chartreuse3","chartreuse4"),labels=c("Klf6","Klf7"))+theme(legend.position = "bottom")
```
Heatmap of Klf67/913 union
```{r}
KLF913<-lfcShrink(dds, contrast = c("Group","Klf913_P20","ScrGFP_P20"),type="ashr")

K913df<-as.data.frame(KLF913)%>%rownames_to_column("ENSID")%>%
  left_join(bmart1,by = c("ENSID"="ensembl_gene_id"))%>%
  filter(chromosome_name%in%c(as.character(1:19),"X","Y"),gene_biotype%in%c("protein_coding"))
K913df_filt<-K913df%>%filter(abs(log2FoldChange)>=lfc,padj<=pval)

#intergene<-K67df10_filt%>%filter(external_gene_name%in%intersect(K67df10_filt$external_gene_name,K913df_filt$external_gene_name))
K913df_filt2<-K913df%>%filter(padj<=pval,abs(log2FoldChange)>=log2(1.5))
K67df10_filt2<-K67df_10%>%filter(padj<=pval,abs(log2FoldChange)>=log2(1.5))
intergene<-K913df%>%filter((ENSID%in%K67df10_filt$ENSID&ENSID%in%K913df_filt2$ENSID)|(ENSID%in%K913df_filt$ENSID&ENSID%in%K67df10_filt2$ENSID))

countmat_rlg2<-countmat_rlg[,c(13:16,1:4,17:20,9:12,21)]
DEGset<-intergene
Candgenes<-c(intersect(DEGset$external_gene_name,axongenes),
                grep(DEGset$external_gene_name,pattern=IonChannels,value=TRUE))

y<-heatmapplus2(countmat_rlg2,target=c("Klf9/13","Klf6/7"),DEGset = intergene,row_names_gp = gpar(fontsize= 10),remix=FALSE,column_split=rep(c("A","B"),each=8),column_gap=unit(1.5,"mm"),returnraw = TRUE)
draw(y,heatmap_legend_side = "bottom")
```


```{r}
K913df_filt2<-K913df%>%filter(padj<=pval)
K67df10_filt2<-K67df_10%>%filter(padj<=pval)
K913_67<-K913df[,c(1,3,6,7)]%>%inner_join(K67df_10[,c(1,3,6,7)],by="ENSID",suffix=c("_Klf913","_Klf67"))%>%mutate(DEG=case_when((ENSID%in%K913df_filt2$ENSID&ENSID%in%K67df10_filt2$ENSID) ~ "Shared Target", ENSID%in%K913df_filt2$ENSID ~ "Klf9/13 Target",ENSID%in%K67df10_filt2$ENSID ~ "Klf6/7 Target", .default = "Not Target"))%>%mutate(Shared=ifelse(DEG=="Shared Target",T,F))%>%dplyr::filter(!external_gene_name_Klf913%in%c("Klf6","Klf7","Klf13","Klf9"))
big<-ggplot(K913_67,aes(x=log2FoldChange_Klf67,y=log2FoldChange_Klf913))+geom_point(aes(col=DEG,alpha=Shared)) + scale_color_manual(values = c("Not Target" = "grey", "Klf9/13 Target" = "firebrick2", "Klf6/7 Target" = "forestgreen", "Shared Target" = "purple"))  + labs(x="log2FoldChange(Klf6/7 v Scramble, P10)",y="log2FoldChange(Klf9/13 v Scramble, P20)",col="")+ geom_vline(xintercept = 0,linetype="dashed")+geom_hline(yintercept = 0,linetype="dashed") + theme_bw()+scale_alpha_manual(values=c(0.4,1))
out<-big+theme(axis.title = element_text(size=12))+stat_smooth(data=subset(K913_67,DEG=="Shared Target"),col="purple",method=lm,fullrange = T)+stat_cor(data=subset(K913_67,DEG=="Shared Target"),col="purple",label.y=4,label.x=1,size=5)+
  scale_y_continuous(limit=c(-6,6))+
  scale_x_continuous(limit=c(-6,6))+
  theme(axis.title =element_text(size=14),axis.text =element_text(size=13))
```

Gene Ontology for Shared targets
```{r}
#Attempt 1: Enrichment (only 1 category)
GOdown<- enrichGO(gene = intergene$ENSID, universe = geneuniv,
                OrgDb = org.Mm.eg.db,
                ont = "ALL",minGSSize = 25,
                pAdjustMethod = "BH",keyType = "ENSEMBL",
                pvalueCutoff  = 0.05,qvalueCutoff = 0.1,
                readable = TRUE)

#Attempt 2: GSEA (sorted by ABSOLUTE PC2)
indx<-order(apply(as.data.frame(assay(rlg)),1,var),decreasing = T)
pcx<-prcomp(t(assay(rlg)[indx,]))
pcgenes<-pcx$rotation
topgenes<-as.data.frame(pcgenes)%>%rownames_to_column("ENSID")%>%arrange(desc(PC2))%>%dplyr::select(ENSID)

gsea_genes<-pcgenes[,2]
gsea_genes<-gsea_genes[!is.na(gsea_genes)]

set.seed(101)
GSEAres<- gseGO(geneList = sort(abs(gsea_genes),decreasing = TRUE),
                OrgDb = org.Mm.eg.db,
                ont="ALL",minGSSize = 30,
                pAdjustMethod = "BH",
                keyType = "ENSEMBL",
                pvalueCutoff = 0.05,
                scoreType="pos",eps = 0,
                nPermSimple=10000)

ridgeplot(simplify(GSEAres,cutoff=0.3))

```

Bar graphs of selected targets
```{r}
quickplotKLF<-function(TPMmatrix,genes=NA){
  TPMmatrix%>%filter(external_gene_name%in%genes)%>%
    dplyr::select(contains("Scr"),contains("Klf"),"external_gene_name")%>%
    pivot_longer(cols = -external_gene_name)%>%mutate(virus=str_remove(name,"_(10)?[A-Z][0-9]?$"),age=ifelse(grepl(.$name,pattern = "10"),"P10","P20"))%>%
    mutate(group=paste(virus,age,sep = " "))%>%
    group_by(external_gene_name,group)%>%
    mutate(mean=mean(value),std=sd(value),
           group=factor(group,levels = c("Klf67 P10","ScrGFP P10","Klf67 P20","ScrGFP P20","Klf913 P20")))%>%
    ggplot(aes(x=external_gene_name,y=mean,fill=group))+
    geom_bar(stat="summary",position="dodge",color="black")+
    geom_errorbar(aes(ymin=mean-std,ymax=mean+std),position=position_dodge(.9),width=0.2)+
    geom_point(aes(y=value),position=position_dodge(width=0.9),show.legend = F)+theme_classic()+
    scale_y_continuous(expand=c(0,NA))+
    labs(x="Gene",y="Normalized expression (TPM)",fill="Target")+scale_fill_manual(values=c("lightgreen","lightblue2","forestgreen","blue4","firebrick2"))+
    theme(legend.position = "right",axis.title =element_text(size=15),axis.text.x =element_text(size=15,vjust = 0.6))
}

TPMmatrix<-read.csv("~/Documents/Counts/KLF_GFP_TPM.csv",row.names = 1)
```

```{r}
DEGset<-K67df10_filt%>%filter(log2FoldChange<=1)
countmat_filt<-countmat_rlg%>%filter(GeneID%in%DEGset$external_gene_name)
scale_counts_filt<-apply(countmat_filt[,-c(1:6,16)],1,FUN=function(x) (x-mean(x))/sd(x))
mat<-as.data.frame(t(scale_counts_filt))%>%rownames_to_column("ENSID")%>%pivot_longer(-ENSID,names_to = "Sample")%>%
  mutate(Celltype=str_remove(Sample," [A-Z]$"))%>%group_by(ENSID,Celltype)%>%summarize(avgexp=mean(value))%>%
  mutate(Celltype=factor(Celltype,levels = c("Scramble P10","Klf9/13 P20","Klf6/7 P10","Scramble P20","Klf6/7 P20")))%>%
  group_by(Celltype)
plotx<-mat%>%ggplot(aes(x=Celltype,y=avgexp))+geom_line(aes(group=ENSID),alpha=0.1)+  stat_summary(fun=mean, group=1, geom="line", colour="blue",size=1)+stat_summary(fun=mean,fun.min=function(x) mean(x)-sd(x),fun.max=function(x) mean(x)+sd(x),geom="errorbar", colour="blue",width=0.2,size=1)+labs(x="gRNA Target", y="Normalzied Expresion")+theme_bw()+theme(axis.title =element_text(size=17),axis.text =element_text(size=15))

DEGset<-K913df_filt%>%filter(log2FoldChange>=1)
countmat_filt<-countmat_rlg%>%filter(GeneID%in%DEGset$external_gene_name)
scale_counts_filt<-apply(countmat_filt[,-c(7:9,16)],1,FUN=function(x) (x-mean(x))/sd(x))
mat<-as.data.frame(t(scale_counts_filt))%>%rownames_to_column("ENSID")%>%pivot_longer(-ENSID,names_to = "Sample")%>%
  mutate(Celltype=str_remove(Sample," [A-Z]$"))%>%group_by(ENSID,Celltype)%>%summarize(avgexp=mean(value))%>%
  mutate(Celltype=factor(Celltype,levels = c("Scramble P10","Klf9/13 P20","Klf6/7 P10","Scramble P20","Klf6/7 P20")))%>%
  group_by(Celltype)
plotx<-mat%>%ggplot(aes(x=Celltype,y=avgexp))+geom_line(aes(group=ENSID),alpha=0.1)+  stat_summary(fun=mean, group=1, geom="line", colour="blue",size=1)+stat_summary(fun=mean,fun.min=function(x) mean(x)-sd(x),fun.max=function(x) mean(x)+sd(x),geom="errorbar", colour="blue",width=0.2,size=1)+labs(x="gRNA Target", y="Normalzied Expresion")+theme_bw()+theme(axis.title =element_text(size=17),axis.text =element_text(size=15))
```

Klf67 Age qPCR (w P5)

```{r}
chk<-fread("~/Documents/qPCR/Klf67913_Age/K67Ageqpcr.csv")
KLF<-chk%>%dplyr::select(SampID,FC_K6_P5,FC_K7_P5,Virus,Age)%>%
  pivot_longer(cols=c(FC_K6_P5,FC_K7_P5),names_to = "Target",values_to = "FC")%>%
  mutate(Target=ifelse(Target=="FC_K6_P5","Klf6","Klf7"))%>%arrange(Target,desc(Virus),desc(Age))%>%
  group_by(Virus,Age,Target)%>%mutate(fcmean=mean(FC))
n<-length(unique(KLF$SampID))
KLF$controlmean<-c(rep(KLF$fcmean[1],n),rep(KLF$fcmean[n+1],n))
KLF<-KLF%>%mutate(fcnorm=FC/controlmean)%>%mutate(normmean=mean(fcnorm),fcsd=sd(fcnorm))%>%mutate(Group=paste(Virus, Age, sep=" "))%>%
  mutate(Group=factor(Group,levels = c("Scr-GFP P5","Klf67-GFP P5","Scr-GFP P10-12","Klf67-GFP P10-12","Scr-GFP P18-20","Klf67-GFP P18-20")))
y<-ggplot(KLF,aes(x=Target,y=normmean,fill=Group))+geom_bar(stat="summary",position="dodge",color="black")+
  geom_point(aes(x=Target,y=FC),position = position_dodge(width=0.9),show.legend = F)+
  geom_errorbar(aes(ymin=normmean-fcsd,ymax=normmean+fcsd,x=Target),position = position_dodge(width=0.9),width=0.2)+
  scale_y_continuous(expand=c(0,0),limit=c(0,1.75))+theme_classic()+scale_fill_brewer(palette = "Paired")+
  labs(x="Measured Gene",y="Relative mRNA Levels",fill="gRNA Target")+theme(legend.position = "right",text=element_text(size=16))+
  geom_hline(yintercept = 1,linetype="dashed")+theme(text=element_text(size=18),legend.position = "right",legend.text = element_text(size=12))

```
Nonsense Reviewer 1 plots
```{r}
Dev913<-K913df_old[,c(1,3,5,6,7)]%>%inner_join(Devdf[,c(1,3,6)],by=c("external_gene_name"="ENSID"),suffix=c("_KLF","_Dev"))%>%
  mutate(DEG=case_when(ENSID%in%K913df_filtold$ENSID,"Klf9/13 Target","Not Target"))%>%dplyr::filter(!external_gene_name%in%c("Klf9","Klf13"))

out<-ggplot(Dev913,aes(x=log2FoldChange_Dev,y=log2FoldChange_KLF))+geom_point(aes(col=DEG,alpha=DEG)) + scale_color_manual(values = c("Not Target" = "black", "Klf9/13 Target" = "firebrick2")) + geom_point(data=subset(Dev913,ENSID%in%K67df10_filt$ENSID),color="forestgreen")+
  labs(x="log2FoldChange(P2 v P30)",y="log2FoldChange(Klf9/13 v Scramble)",col="")+geom_vline(xintercept = 0,linetype="dashed")+geom_hline(yintercept = 1,linetype="dashed")+
  geom_hline(yintercept = -1,linetype="dashed") + theme_bw()+
  stat_smooth(data=subset(Dev913,DEG=="Klf9/13 Target"),col = "firebrick2",method=lm,fullrange = T) + stat_smooth(data=subset(Dev913,ENSID%in%K67df10_filt$ENSID),color="forestgreen",method=lm,fullrange=T)+stat_cor(data=subset(Dev913,ENSID%in%K67df10_filt$ENSID),col="forestgreen",label.x=4,label.y=3,size=5,p.accuracy = 0.0000001)+
  stat_cor(data=subset(Dev913,DEG=="Klf9/13 Target"),col="firebrick2",label.x=4,label.y=3.5,size=5)+scale_alpha_manual(values=c(1,0.4))+
  theme(axis.title =element_text(size=16),axis.text =element_text(size=13))+ geom_text_repel(data=subset(Dev913,external_gene_name%in%intersect(KLFbestgenes,intergene$external_gene_name)),                                                                                        aes(x=log2FoldChange_Dev,y=log2FoldChange_KLF,label=external_gene_name),color="grey20",max.overlaps = 25,force=10,fontface="italic")

Dev67<-K67df_10[,c(1,3,5,6,7)]%>%inner_join(Devdf[,c(1,3,6)],by=c("external_gene_name"="ENSID"),suffix=c("_KLF","_Dev"))%>%
  mutate(DEG=ifelse(ENSID%in%K67df10_filt$ENSID,"Klf6/7 Target","Not Target"))%>%dplyr::filter(!external_gene_name%in%c("Klf6","Klf7","Tmem150b"))

out<-ggplot(Dev67,aes(x=log2FoldChange_Dev,y=log2FoldChange_KLF))+geom_point(aes(col=DEG,alpha=DEG)) + scale_color_manual(values = c("Not Target" = "black", "Klf6/7 Target" = "forestgreen")) + geom_point(data=subset(Dev67,ENSID%in%K913df_filt$ENSID),color="firebrick2")+ 
  labs(x="log2FoldChange(P2 v P30)",y="log2FoldChange(Klf6/7 v Scramble, P10)",col="")+geom_vline(xintercept = 0,linetype="dashed")+geom_hline(yintercept = 1,linetype="dashed")+
  geom_hline(yintercept = -1,linetype="dashed") + theme_bw() +stat_smooth(data=subset(Dev67,ENSID%in%K913df_filt$ENSID),color="firebrick2",method=lm,fullrange=T)+stat_cor(data=subset(Dev67,ENSID%in%K913df_filt$ENSID),col="firebrick2",label.x=-10,label.y=2,size=5,p.accuracy = 0.0000001)+
  stat_smooth(data=subset(Dev67,DEG=="Klf6/7 Target"),col = "forestgreen",method=lm,fullrange = T) + 
  stat_cor(data=subset(Dev67,DEG=="Klf6/7 Target"),col="forestgreen",label.x=-10,label.y=2.5,size=5)+
  theme(axis.title =element_text(size=16),axis.text =element_text(size=13))+
  scale_alpha_manual(values=c(1,0.4))+geom_text_repel(data=subset(Dev67,external_gene_name%in%intersect(KLFbestgenes,intergene$external_gene_name)),                                                                                        aes(x=log2FoldChange_Dev,y=log2FoldChange_KLF,label=external_gene_name),color="grey20",max.overlaps = 25,force=10,fontface="italic")
