---
title: "R Notebook"
output: html_notebook
---

```{r}
library(DESeq2)
library(apeglm)
library("RColorBrewer")
library(pheatmap)
library(ComplexHeatmap)
library(tidyverse)
library(GenomicRanges)
library(data.table)
library(biomaRt)
```

Run first ALWAYS
```{r}
setwd("~/Documents/DevATAC/")
ensembl<-useMart(host="https://may2024.archive.ensembl.org",biomart = "ENSEMBL_MART_ENSEMBL",dataset = "mmusculus_gene_ensembl")
bmart1<-getBM(attributes = c("ensembl_gene_id", "external_gene_name","chromosome_name","gene_biotype"), mart = ensembl)
bfilt<-bmart1%>%filter(chromosome_name%in%c(as.character(1:19),"X","Y"),gene_biotype%in%c("protein_coding"))
```



Alt input: STAR-mapped samples (mm10 or GENCODE M27)
```{r}
countmat<-read.csv("~/Documents/DevATAC/P2P30_RNA/p2p30_mm10_counts.csv")%>%mutate(X=make.unique(X))%>%column_to_rownames("X")
countmat<-countmat[,c(1:4,13:16,21:28)]

TPMmat<-read.csv("P2P30_RNA/p2p30_mm10_TPM_s1.csv",row.names = 1)
#rownames(TPMmat)<-str_remove(rownames(TPMmat),pattern = "\\.[0-9]*$")
colmat<-data.frame(Layer=ifelse(grepl(names(countmat),pattern = "L4"),"L4","L6"),
                   Age=str_extract(names(countmat),pattern = "p[0-9]+"),row.names = names(countmat))%>%
  mutate(Age=replace_na(Age,replace="p30"))%>%
  mutate(Group=paste(Age,Layer,sep="_"))
ddsTxi<-DESeqDataSetFromMatrix(countmat,
                               colData = colmat,
                               design = ~ Group)
```

Prefiltering for DESeq2
```{r}
keep<- apply(counts(ddsTxi),1,function(x) any(x>10))

ddsTxi <- ddsTxi[keep,]

ddsTxi<-DESeq(ddsTxi)
rlg<-as.data.frame(assay(rlog(ddsTxi,blind = FALSE)))
#vstdf<-as.data.frame(assay(vst(ddsTxi,blind = FALSE)))
```


```{r}
tst<-plotPCA(rlog(ddsTxi,blind = FALSE),intgroup=c("Group"),ntop=1000)+theme_bw()+labs(color= "Group")
tst$data<-tst$data%>%mutate(PC1=PC1*-1)
out<-tst+scale_fill_manual(values=c("salmon","lightblue3","red","blue"),name="Group",labels=c("P2 Layer 4 ","P2 Layer 6","P30 Layer 4","P30 Layer 6"))+
  theme(axis.title =element_text(size=14),legend.position = "left")+
  geom_point(aes(fill=group),color="black",pch=21,size=6)

```

Get Age-related DEGs for each layer *separately  

```{r}
L4shrink<-lfcShrink(ddsTxi,contrast = c("Group","p30_L4","p2_L4"),type = "ashr")
L6shrink<-lfcShrink(ddsTxi,contrast = c("Group","p30_L6","p2_L6"),type = "ashr")


#p2shrink<-lfcShrink(ddsTxi,contrast = c("Group","p2_L4","p2_L6"),type = "ashr")
#p30shrink<-lfcShrink(ddsTxi,contrast = c("Group","p30_L4","p30_L6"),type = "ashr")
```

**New and Improved way to get shared vs. all DEGs
```{r}
padj<-0.01
lfc<-1
basemean<-20
TPMfilt<-TPMmat$external_gene_name[apply(TPMmat[,-17],1,max)>=10]
L4df<-as.data.frame(L4shrink)[,-c(3,4)]%>%rownames_to_column("ENSID")
L6df<-as.data.frame(L6shrink)[,-c(3,4)]%>%rownames_to_column("ENSID")
```

```{r}
#If starting from mm10 (refseq) mapping 
L4_6df<-left_join(L4df,L6df,by="ENSID",suffix=c("L4","L6"))
L4_6df_filt<-L4_6df%>%filter(ENSID%in%bfilt$external_gene_name,ENSID%in%TPMfilt)
L4<-L4df%>%filter(abs(log2FoldChange)>lfc,padj<=padj,baseMean>basemean)%>%
  left_join(bmart1,by=c("ENSID"="external_gene_name"))%>%filter(ENSID%in%bfilt$external_gene_name)
L6<-L6df%>%filter(abs(log2FoldChange)>=lfc,padj<=padj,baseMean>basemean)%>%
  left_join(bmart1,by=c("ENSID"="external_gene_name"))%>%filter(ENSID%in%bfilt$external_gene_name)
L4_6df_all<-L4_6df%>%filter((abs(log2FoldChangeL4)>=lfc&padjL4<=padj)|(abs(log2FoldChangeL6)>=lfc&padjL6<=padj),(baseMeanL4>basemean | baseMeanL6>basemean))%>%
left_join(bmart1,by=c("ENSID"="external_gene_name"))%>%filter(ENSID%in%bfilt$external_gene_name,ENSID%in%TPMfilt)
L4_6df_shared<-L4_6df%>%filter(abs(log2FoldChangeL4)>=lfc,padjL4<=padj,abs(log2FoldChangeL6)>=lfc,padjL6<=padj,(baseMeanL4>basemean | baseMeanL6>basemean))%>%
  left_join(bmart1,by=c("ENSID"="external_gene_name"))%>%filter(ENSID%in%bfilt$external_gene_name)
```

ALL Genes Heatmap (uses mean expression of each sample)
```{r}
library(ComplexHeatmap)
#Prefiler and summarize data
#inmat<-as.data.frame(assay(rlg_batch_filt))
inmat<-rlg
#inmat<-log2(TPMmat[,-17]+1)

countmat_filt<-inmat%>%rownames_to_column("ENSID")%>%filter(ENSID%in%L4_6df_all$ENSID)%>%
  pivot_longer(-ENSID)%>%mutate(Layer=ifelse(grepl(name,pattern = "L4"),"Layer 4","Layer 6"),Age=ifelse(grepl(name,pattern="_p[2-3]_"),"P2","P30"),
                                Group=paste(Age,Layer,sep=" "))%>%
  group_by(Group,ENSID)%>%summarise(mean=mean(value))%>%
  pivot_wider(names_from = Group,values_from=mean)%>%column_to_rownames("ENSID")

scale_counts_filt<-apply(countmat_filt,1,FUN=function(x) (x-mean(x))/sd(x))
```

K means clustering, dendrogram tree cut method
```{r}
#Find optimal cluster number using Gap statistic (takes a second)
#gapstat<-clusGap(t(scale_counts_filt2),FUNcluster = kmeans,K.max = 10,iter.max=30,nstart=100)
#Found optimum 
km<-6

set.seed(123)
genekm<-kmeans(t(scale_counts_filt),centers=km,iter.max=30,nstart=100)
mat<-Heatmap(t(scale_counts_filt),
             cluster_columns = F,clustering_method_rows = "average",clustering_distance_rows = "pearson",show_row_names = F,
             col = colorRampPalette(c("yellow","black","magenta"))(100),name="Normalized Expression",
             column_names_rot=45,heatmap_legend_param = list(direction="horizontal"),
             row_split = factor(as.numeric(genekm$cluster),levels=c(1,6,2,4,5,3)),show_parent_dend_line = FALSE)

#Run to see heatmap and perform clustering
kmat<-draw(mat,heatmap_legend_side = "bottom")
```

```{r}
#retrieve results and group genes
klist<-row_order(kmat)
temp<-data.frame(RowID=do.call(c,klist),Cluster=rep(names(klist),sapply(klist,length)))

temp2<-as.data.frame(t(scale_counts_filt)[temp$RowID,])%>%mutate(Group=temp$Cluster)%>%rownames_to_column("ENSID")%>%
  pivot_longer(-c(ENSID,Group),names_to = "Sample")%>%
  group_by(Group,Sample)%>%mutate(avgexp=mean(value))

temp<-temp%>%mutate(GeneID=colnames(scale_counts_filt)[temp$RowID])

#Plot clusters
plotx<-temp2%>%ggplot(aes(x=Sample,y=avgexp,col=Group,group=Group))+geom_line()+geom_line(aes(y=value,group=ENSID),alpha=0.01)+
  stat_summary(mapping = aes(y=value),fun=mean,group=1,fun.min=function(x) mean(x)-sd(x),fun.max=function(x) mean(x)+sd(x),geom="errorbar",width=0.2)+
  labs(x="Group", y="Normalzied Expresion")+facet_wrap(~Group,ncol=2)+theme_bw()+theme(axis.text.x = element_text(angle = 45,vjust=0.5))+
  labs(x=NULL,col="Cluster")+scale_color_brewer(palette = "Dark2")
```

Same as above on full dataset (not averages)
```{r}
countmat_filt2<-inmat%>%rownames_to_column("ENSID")%>%filter(ENSID%in%intergene$external_gene_name_Klf67)%>%
  column_to_rownames("ENSID")
countmat_filt2<-countmat_filt2[,c(5:12,1:4,13:16)]
names(countmat_filt2)<-paste(paste(rep(c("P2","P30"),each=8),rep(c("Layer 4","Layer 6"),each=4)),c("A","B","C","D"))


scale_counts_filt2<-apply(countmat_filt2,1,FUN=function(x) (x-mean(x))/sd(x))
km<-6

set.seed(123)
kmout<-kmeans(t(scale_counts_filt2),centers=6,iter.max=30,nstart=100)

col<-c("lightskyblue","indianred","navy","red3")
collabs=HeatmapAnnotation(Group=str_remove(rownames(scale_counts_filt2),pattern = ' [A-Z]+$'),
                          col=list(Group=c("P2 Layer 4"=col[1],"P2 Layer 6"=col[2],"P30 Layer 4"=col[3],"P30 Layer 6"=col[4])), show_annotation_name=FALSE)
mat<-Heatmap(t(scale_counts_filt2),
             cluster_columns = F,clustering_method_rows = "average",clustering_distance_rows = "pearson",show_row_names = F,
             col = colorRampPalette(c("magenta3","black","green3"))(100),name="Normalized Expression",
             column_names_rot=45,heatmap_legend_param = list(direction="horizontal"),top_annotation = collabs,
             row_split = factor(as.numeric(kmout$cluster),levels=c(2,4,5,6,1,3)),show_parent_dend_line = FALSE,show_column_names = F,border = NULL)

#Run to see heatmap and perform clustering
kmat<-draw(mat,heatmap_legend_side = "bottom")
```

GO for shared clusters
```{r}
library(clusterProfiler)
library(org.Mm.eg.db)
upshared<-names(genekm$cluster[genekm$cluster==1])
downshared<-names(genekm$cluster[genekm$cluster==3])
L46downshared<-L4_6df_all%>%filter(ENSID%in%downshared)
L46upshared<-L4_6df_all%>%filter(ENSID%in%upshared)
geneuniv<-unique(L4_6df$ENSID)

GOup<- enrichGO(gene = upshared, universe = geneuniv,
                OrgDb = org.Mm.eg.db,
                ont = "ALL",minGSSize = 25,
                pAdjustMethod = "BH",keyType = "SYMBOL",
                pvalueCutoff  = 0.05, qvalueCutoff = 0.05,
                readable = TRUE)
GOdown<- enrichGO(gene = downshared, universe = geneuniv,
                OrgDb = org.Mm.eg.db,
                ont = "ALL",minGSSize = 25,
                pAdjustMethod = "BH",keyType = "SYMBOL",
                pvalueCutoff  = 0.05,qvalueCutoff = 0.05,
                readable = TRUE)

options(enrichplot.colours=c("red","blue"))
GOup@result<-GOup@result[!str_detect(GOup@result$Description,"mito"),]
#barplot(simplify(GOup,0.7),showCategory=10)+scale_y_discrete(position="right")
#barplot(simplify(GOdown,0.7),showCategory=10)+scale_y_discrete(position="right")

#get some interesting functional genes for labeling
topgenes<-GOup@result%>%
  filter(Description%in%c("regulation of monoatomic ion transport","neuron to neuron synapse","regulation of vesicle-mediated transport"))
topgenes<-unique(str_split(paste(topgenes$geneID,collapse = "/"),"/")[[1]])

topgenes2<-GOdown@result%>%
  filter(Description%in%c("axon guidance","regulation of neuron projection development","regulation of nervous system development"))
topgenes2<-unique(str_split(paste(topgenes2$geneID,collapse = "/"),"/")[[1]])

topgenes<-c(topgenes,topgenes2)
```

```{r}
simplify(GOup,0.7)@result%>%as.data.frame()%>%mutate(yq=-log10(qvalue))%>%arrange(desc(yq))%>%head(10)%>%ggplot(aes(y=yq,x=reorder(Description,yq)))+geom_col(fill="purple4",col="black")+coord_flip()+theme_classic()+labs(y=expression(-log[10](q.val)),x="")+theme(axis.text.y = element_text(size=16,face="bold"),text=element_text(size=18))+ scale_y_continuous(expand = c(0,0),limits = c(0,5))+scale_x_discrete(position = "top")

simplify(GOdown,0.7)@result%>%as.data.frame()%>%mutate(yq=-log10(qvalue))%>%arrange(desc(yq))%>%head(10)%>%ggplot(aes(y=yq,x=reorder(Description,yq)))+geom_col(fill="mediumorchid",col="black")+coord_flip()+theme_classic()+labs(y=expression(-log[10](q.val)),x="")+theme(axis.text.y = element_text(size=16,face="bold"),text=element_text(size=18))+ scale_y_continuous(expand = c(0,0),limits = c(0,5))+scale_x_discrete(position = "top")
```


Make GRanges for overlap testing
```{r}
mart<-useMart(host="https://jan2020.archive.ensembl.org",biomart = "ENSEMBL_MART_ENSEMBL",dataset = "mmusculus_gene_ensembl")
bmart2<-getBM(attributes = c("ensembl_gene_id","ensembl_transcript_id","transcript_start",
"transcript_end","chromosome_name","strand","transcript_biotype","transcript_gencode_basic"), mart = mart)

GRangefrombmart<-function(bmart,filter=TRUE){
  bmart<-bmart%>%filter(chromosome_name%in%c(as.character(1:19),"X","Y"))
  if(filter==TRUE){bmart<-bmart%>%filter(transcript_biotype=="protein_coding",
                                   transcript_gencode_basic=="GENCODE basic")}
  out<-GRanges(seqnames=paste0("chr",bmart$chromosome_name),
          ranges=IRanges(start=bmart$transcript_start,end=bmart$transcript_end),
          strand=ifelse(bmart$strand==1,"+","-"),
          mcols=data.frame(ENSID=bmart$ensembl_transcript_id,GeneID=bmart$ensembl_gene_id))
  outlist<-split(out,out$mcols.GeneID)
  outlist<-range(outlist)
  outplus<-unlist(outlist)
  return(sort(outplus))
}

TSSfrombmart<-function(bmart,extend=c(100000,100000),reduce=TRUE,filter=TRUE){
  bmart<-bmart%>%filter(chromosome_name%in%c(as.character(1:19),"X","Y"))
    if(filter==TRUE){bmart<-bmart%>%filter(transcript_biotype=="protein_coding",
                                   transcript_gencode_basic=="GENCODE basic")}
  out<-GRanges(seqnames=paste0("chr",bmart$chromosome_name),
          ranges=IRanges(start=bmart$transcript_start,end=bmart$transcript_end),
          strand=ifelse(bmart$strand==1,"+","-"),
          mcols=data.frame(ENSID=bmart$ensembl_transcript_id,GeneID=bmart$ensembl_gene_id))
  outplus<-promoters(out,upstream = extend[1],downstream = extend[2])
  if(reduce==TRUE){outlist<-split(outplus,outplus$mcols.GeneID)
  outlist<-GenomicRanges::reduce(outlist)
  outplus<-unlist(outlist)
  }
  return(sort(outplus))
}

#Get genes and TSS windows as GRanges
ensgenes<-GRangefrombmart(bmart2)
TSSgenes<-TSSfrombmart(bmart2,extend=c(1000,200))
```

New GRanges maker (RIP Biomart archives)
```{r}
library(rtracklayer)
library(ensembldb)
#gff<-import("Documents/Mus_musculus.GRCm38.102.gtf.gz")
#mcols(gff[,c(12,16,19,20)])
goodtx<-as.data.frame(mcols(gff[,c(2,12,16,19,20)]))%>%
  filter(type=="transcript",transcript_biotype=="protein_coding",tag=="basic")

#enspath<-ensembldb::ensDbFromGtf("Documents/Mus_musculus.GRCm38.102.gtf.gz")
ensdb102<-EnsDb(enspath)
tx <- transcripts(ensdb102, columns = c("tx_id", "gene_id"))
#Filter out extra chromosommes
txgood<-tx[tx@seqnames%in%c(as.character(1:19),"X","Y")]

#Filter by protein-coding + GENCODE BASIC annotation
txgood<-txgood[txgood$tx_id%in%goodtx$transcript_id]

#Organize as genes
templist<-split(txgood,txgood$gene_id)
templist<-range(templist)
ensgenes<-unlist(templist)
ensgenes<-sort(txgood_plus)

#Get TSS
txgood_plus<-promoters(txgood,upstream = 1000,downstream = 200)
templist<-split(txgood_plus,txgood_plus$gene_id)
templist<-GenomicRanges::reduce(templist)
TSSgenes<-unlist(templist)
TSSgenes<-sort(TSSgenes)
```


```{r}
library(ggrepel)
volcanoooTF<-function(infile,pval=0.01,motifs=NA,allpostext=F,JASPARID=T,onlypos=F,...){
  if(JASPARID){infile$TFID<-str_split(infile$TFID,pattern = "_",simplify = T)[,1]}
  if(!is.na(motifs)){infile<-mutate(infile,isMotif=ifelse(grepl(TFID,pattern=motifs,ignore.case=T),T,F))
  } else {infile<-mutate(infile,isMotif=F)}
  if(onlypos){infile<-mutate(infile,isMotif=ifelse(padj>pval,F,isMotif))}
  temp<-infile%>%filter(!is.nan(padj))%>%
                          mutate(Effect=case_when(Enrichment>1 & padj<pval ~"Enriched(Test)",
                             Enrichment<1 & padj<pval ~ "Enriched(Control)",
                             TRUE ~ "No Change"))
  temptab<-table(temp$Effect)
  fig1<-temp%>%ggplot(aes(x=Enrichment,y=-log10(padj)))+geom_point(aes(col=Effect))+
  scale_color_manual(breaks = c("Enriched(Test)","Enriched(Control)","No Change"),
                     values=ggplot2::alpha(c("red", "blue","gray30"),0.5))+
  geom_point(data=subset(temp,isMotif==T),shape=21,aes(x=Enrichment,y=-log10(padj),fill=Effect),
             color="black",stroke=1.2)+
  scale_fill_manual(breaks = c("Enriched(Test)","Enriched(Control)","No Change"),
                     values=c("red", "blue","gray30"))+
    geom_text_repel(data=subset(temp,isMotif==T),
                    aes(x=Enrichment,y=-log10(padj),label=TFID),...)+
  ylab(expression(-log[10](p.val)))+
    xlab("Enrichment")+
    theme_classic()+
    geom_hline(yintercept = -log10(pval),linetype="dashed")+
     geom_vline(xintercept = 1,linetype="dashed")+
  theme(axis.title=element_text(size=15),axis.text = element_text(size=15),legend.position = "bottom",
        legend.text = element_text(size=15),legend.title = element_text(size=18))+guides(fill="none")
  if(allpostext==T){
    fig1<- fig1 + geom_text_repel(data=subset(temp,Effect=="Enriched(Test)"),
                                  aes(x=Enrichment,y=-log10(padj),label=TFID),...)
  }
  return(fig1)
}

contingency_mat<-function(col,hitmat=temp){
  pos_test<-sum(col[hitmat])
  neg_test<-length(col[hitmat])-pos_test
  pos_cont<-sum(col[!hitmat])
  neg_cont<-length(col[!hitmat])-pos_cont
  out<-matrix(c(pos_test,neg_test,pos_cont,neg_cont),nrow=2,ncol=2,byrow=T)
  return(out)
}
```


Load Subset Peak files
```{r}
load("P2P30_ATAC/Sacha/p2p30Peaks_rawcounts.RDa",verbose = T)
devpeakGR<-devcountGR
mcols(devpeakGR)<-NULL
```

```{r}
library(ChIPseeker)
library(TxDb.Mmusculus.UCSC.mm10.ensGene)
txdb<-TxDb.Mmusculus.UCSC.mm10.ensGene
AllPeakAnno<-annotatePeak(devpeakGR,TxDb=txdb,tssRegion = c(-1000,200),flankDistance = 10000)

AllPeakAnno@anno<-as.data.frame(AllPeakAnno@anno)%>%
  mutate(fixedGene=case_when(grepl(annotation,pattern = "ENS")
                                              ~str_extract(annotation,pattern = "ENSMUSG[0-9]*"),
                                              TRUE ~ str_extract(geneId,pattern = "ENSMUSG[0-9]*")))%>%
  mutate(annotation=str_split_fixed(annotation,pattern = " \\(",n = 2)[,1])%>%
  left_join(bmart1[,-3],by=c("fixedGene"="ensembl_gene_id"))%>%GRanges()

IntpeaksFULL<-AllPeakAnno@anno[AllPeakAnno@anno$annotation=="Intron"]

PropeaksFULL<-AllPeakAnno@anno[AllPeakAnno@anno$annotation=="Promoter"]

IGpeaksFULL<-AllPeakAnno@anno[AllPeakAnno@anno$annotation=="Distal Intergenic"]

AllPeaksFULL<-AllPeakAnno@anno
```

```{r}
peakcounts<-as.data.frame(mcols(devcountGR))
rownames(peakcounts)<-paste(seqnames(devcountGR),start(devcountGR),sep="_")
peakcol<-data.frame(row.names = names(peakcounts),Layer=str_extract(names(peakcounts),pattern = "^L[0-9]+"),
                    Age=ifelse(grepl(names(peakcounts),pattern="p2"),"P2","P30"))%>%
                    mutate(Group=paste(Age,Layer,sep="_"))

ddsATAC<-DESeqDataSetFromMatrix(round(peakcounts),
                               colData = peakcol,
                               design = ~ Group)

keep<- apply(counts(ddsATAC),1,function(x) any(x>50))

ddsATAC <- ddsATAC[keep,]

ddsATAC<-DESeq(ddsATAC)

rlgATAC<-rlog(ddsATAC,blind = FALSE)
#rlgATAC from ATAC
tst<-plotPCA(rlgATAC,intgroup=c("Group"),ntop=1000)+theme_bw()+labs(color= "Group")
tst$data<-tst$data%>%mutate(PC1=PC1*-1)
out<-tst+ theme(aspect.ratio=0.8)+scale_fill_manual(values=c("salmon","lightblue3","red","blue"),name="Group",labels=c("P2 Layer 4 ","P2 Layer 6","P30 Layer 4","P30 Layer 6"))+
  theme(axis.title =element_text(size=14),legend.position = "left")+
  geom_point(aes(fill=group),color="black",pch=21,size=6)
```


```{r}
library(DiffBind)
library(stringr)
#Load in sample matrix
samples=read.csv("~/Documents/DevDiff/Diffbind_p2p30_good200.csv")
blacklist<-rtracklayer::import("~/Documents/DevDiff/mm10blacklistv2.bed")

#Get Preset Peak set
load("~/Documents/DevDiff/p2p30/p2p30PeaksSN_rawcounts.RDa",verbose = T)
mcols(devcountGR)<-NULL
seqlevelsStyle(devcountGR)<-"NCBI"
PeakGR<-subsetByOverlaps(devcountGR,blacklist,invert = TRUE)


dx<-dba(sampleSheet = samples)
dx<-dba.blacklist(dx,blacklist=blacklist,greylist = FALSE)

dc<-dba.count(dx,summits = 100,filter = 10)

#dc<-dba.count(dx,peaks = PeakGR,summits=FALSE,bParallel=TRUE,filter=10)

dc$merged[,1]<-dc$peaks[[1]][,1]
testGR<-GRanges(as.data.frame(dc$merged))
#rtracklayer::export(testGR,"~/Documents/DevDiff/DiffBind_out/DBindmerge_200_filt10.bed")
#save(dc,file = "~/Documents/DevDiff/DiffBind_out/DBindmerge_200_filt10.RDa")
```

```{r}
load("~/Documents/DevATAC//DiffBind/DBindmerge_200_filt10.rda")
dc_norm<-dba.normalize(dc,method = DBA_DESEQ2,library="RiP",normalize="TMM")
allpeaks<-dba.peakset(dc_norm,bRetrieve = T)
seqlevelsStyle(allpeaks)<-"UCSC"

dcx_L4<-dba.contrast(dc_norm,contrast = c("Factor","L4_P30","L4_P2"),minMembers = 2)
dca_out_L4<-dba.analyze(dcx_L4)
dcaGR_L4<-dba.report(dca_out_L4,fold=1,th=0.01)
seqlevelsStyle(dcaGR_L4)<-"UCSC"
                                           
dcx_L6<-dba.contrast(dc_norm,contrast = c("Factor","L6_P30","L6_P2"),minMembers = 2)
dca_out_L6<-dba.analyze(dcx_L6)
dcaGR_L6<-dba.report(dca_out_L6,fold=1,th=0.01)
seqlevelsStyle(dcaGR_L6)<-"UCSC"

names(mcols(dcaGR_L4))[4:6]<-paste0(names(mcols(dcaGR_L4)),"_L4")[4:6]
names(mcols(dcaGR_L6))[4:6]<-paste0(names(mcols(dcaGR_L6)),"_L6")[4:6]
#as.data.frame(mcols(dcaGR_L4))%>%mutate(Change=ifelse(Fold_L4>0,"Up","Down"))%>%group_by(Change)%>%summarise(n())
#as.data.frame(mcols(dcaGR_L6))%>%mutate(Change=ifelse(Fold_L6>0,"Up","Down"))%>%group_by(Change)%>%summarise(n())
                                        
L46allGR<-unique(c(dcaGR_L4,dcaGR_L6))

L46sharedGR<-sort(subsetByOverlaps(dcaGR_L4,dcaGR_L6,type="equal"))
L6tempGR<-sort(subsetByOverlaps(dcaGR_L6,dcaGR_L4,type="equal"))
mcols(L46sharedGR)<-cbind(mcols(L46sharedGR),mcols(L6tempGR)[,-1])

as.data.frame(mcols(L46sharedGR))%>%mutate(L6=ifelse(Fold_L6>0,"L6 Up","L6 Down"),L4=ifelse(Fold_L4>0,"L4 Up","L4 Down"))%>%group_by(L6,L4)%>%summarise(n())
```

```{r}
library(ChIPseeker)
library(TxDb.Mmusculus.UCSC.mm10.ensGene)
txdb<-TxDb.Mmusculus.UCSC.mm10.ensGene
peaksin<-allpeaks
mcols(peaksin)<-NULL
AllPeakAnno<-annotatePeak(peaksin,TxDb=txdb,tssRegion = c(-1000,200),flankDistance = 10000)
#AllPeakAnno2<-annotatePeak(L46allGR,TxDb=txdb,tssRegion = c(-1000,200),flankDistance = 10000)
#L4Anno<-annotatePeak(dcaGR_L4,TxDb=txdb,tssRegion = c(-1000,200),flankDistance = 10000)
#L6Anno<-annotatePeak(dcaGR_L6,TxDb=txdb,tssRegion = c(-1000,200),flankDistance = 10000)
#SharedAnno<-annotatePeak(L46sharedGR,TxDb=txdb,tssRegion = c(-1000,200),flankDistance = 10000)


AllPeakAnno@anno<-as.data.frame(AllPeakAnno@anno)%>%mutate(fixedGene=case_when(grepl(annotation,pattern = "ENS")
                                              ~str_extract(annotation,pattern = "ENSMUSG[0-9]*"),
                                              TRUE ~ str_extract(geneId,pattern = "ENSMUSG[0-9]*")))%>%
  mutate(annotation=str_split_fixed(annotation,pattern = " \\(",n = 2)[,1])%>%
  left_join(bmart1[,-3],by=c("fixedGene"="ensembl_gene_id"))%>%GRanges()

Intpeaks<-AllPeakAnno@anno[AllPeakAnno@anno$annotation=="Intron"]

Propeaks<-AllPeakAnno@anno[AllPeakAnno@anno$annotation=="Promoter"]

IGpeaks<-AllPeakAnno@anno[AllPeakAnno@anno$annotation=="Distal Intergenic"]

AllPeaks<-AllPeakAnno@anno
```

```{r}
Annodf<-data.frame(f1=AllPeakAnno@annoStat[,1],All=AllPeakAnnoFULL@annoStat[,2],DAR=AllPeakAnno2@annoStat[,2],L4=L4Anno@annoStat[,2],L6=L6Anno@annoStat[,2],Shared=SharedAnno@annoStat[,2])%>%
  mutate(Feature=case_when(str_detect(f1,"Intron") ~ "Intron", str_detect(f1,"Downstream|Distal") ~ "Intergenic", 
                            str_detect(f1,"Promoter") ~ "Promoter", TRUE ~ "Exon"))%>%
  dplyr::select(-f1)%>%pivot_longer(-Feature)%>%
  group_by(name,Feature)%>%summarise(Total=sum(value))%>%
  mutate(Feature=factor(Feature,levels=c("Promoter","Intron","Intergenic","Exon")))

statplot<- Annodf%>%ggplot(aes(fill=Feature,x=name,y=Total))+geom_bar(position="stack",stat="identity")+labs(x="Group",y="% Peaks",fill="Genomic Feature")+scale_fill_brewer(palette = "Set1")+theme_minimal()+
  theme(axis.text = element_text(size=12),axis.title = element_text(size=15),legend.position = "bottom")

```

```{r}
peaksin<-devpeakGR
load("~/Documents/DevATAC/p2p30PeaksSN_JASPAR2022CORE_FIMO.rda")
#load("~/Documents/DevATAC/p2p30PeaksSN_JASPAR2022root_FIMO.rda")
testout<-sapply(listy,function(x) countOverlaps(peaksin,x))
mcols(peaksin)<-testout
```

```{r}
peaksinPro<-sort(Propeaks)
testout<-sapply(listy,function(x) countOverlaps(peaksinPro,x))
testout[testout>0]<-1
mcols(peaksinPro)<-testout
peaksinPro$fixedGene<-sort(Propeaks)$fixedGene

dcaPro_P2P30<-subsetByOverlaps(dcaGR_P2P30,Propeaks,type="equal")

mcols(peaksinPro)<-cbind(mcols(sort(dcaPro_P2P30)),mcols(peaksinPro))

peaksinProdf<-as.data.frame(peaksinPro)

DEGset<-filter(K67df10_filt,log2FoldChange<=-1,padj<=0.05)

peaksinProdf%>%mutate(DEG=ifelse(fixedGene%in%DEGset$ENSID,T,F))%>%ggplot(aes(x=Fold,col=DEG))+stat_ecdf()
peaksinProdf%>%mutate(DEG=ifelse(fixedGene%in%DEGset$ENSID,T,F))%>%ggplot(aes(x=Fold,col=DEG))+geom_density()
```


```{r}
downshared<-names(genekm$cluster[genekm$cluster==3])
upshared<-names(genekm$cluster[genekm$cluster==1])

geneset<-upshared
#For Misc. Regulatory Regions
BGset<-PropeaksFULL

#ovtest<-findOverlaps(BGset,peaksin,minoverlap = 100)
ovtest<-findOverlaps(BGset,peaksin,type = "equal")
BGmotifs<-peaksin[ovtest@to]
BGmotifs$geneID<-BGset$external_gene_name[ovtest@from]

#136 for clutser, 837 for all
testout_plus<-mcols(BGmotifs[,-837])

listx<-list()
for(i in seq(ncol(testout_plus))){
  listx[[i]]<-testout_plus[,i]>0
}

temp<-BGmotifs$geneID%in%geneset


#Chitest_pos<-data.frame(Enrichment=sapply(contabs,FUN=function(x) (x[1,1]/sum(x[1,]))/(x[2,1]/sum(x[2,]))))%>%
#  mutate(TFID=names(contabs),Pval=sapply(contabs,FUN=function(x) chisq.test(x)$p.value))%>%
#  mutate(padj=p.adjust(Pval,"BH"))

contabs<-lapply(listx,FUN=contingency_mat)
names(contabs)<-names(listy)

Fishtest<-lapply(contabs,FUN=function(x) fisher.test(x))
Fishtest_pos<-data.frame(Enrichment=sapply(Fishtest,function(x) x$estimate),
                         Pval=sapply(Fishtest,function(x) x$p.value))%>%
                           mutate(TFID=names(contabs),padj=p.adjust(Pval,method="BH"))

volcanoooTF(Fishtest_pos,motifs="Klf|Sox|Sp",onlypos = T)
volcanoooTF(Fishtest_pos,motifs="Cre|Tcf|fos|Jun",onlypos = T)
#volcanoooTF(Fishtest_pos,JASPARID = F,allpostext = T)
```


```{r}
DARAnno<-subsetByOverlaps(AllPeaks,L46allGR,type="equal")

IntDAR<-subsetByOverlaps(Intpeaks,L46allGR,type="equal")

ProDAR<-subsetByOverlaps(Propeaks,L46allGR,type="equal")

IGDAR<-subsetByOverlaps(IGpeaks,L46allGR,type="equal")
```

```{r}
dc_norm2<-dba.normalize(dc,method = DBA_DESEQ2,library="RiP",normalize="lib")
allpeaks<-dba.peakset(dc_norm2,bRetrieve = T)
seqlevelsStyle(allpeaks)<-"UCSC"

inmat<-as.data.frame(mcols(subsetByOverlaps(allpeaks,ProDAR,type="equal")))

countmat_filt<-inmat%>%rownames_to_column("PeakID")%>%
    pivot_longer(-PeakID)%>%mutate(Layer=ifelse(grepl(name,pattern = "L4"),"Layer 4","Layer 6"),Age=ifelse(grepl(name,pattern="p2"),"P2","P30"),
                                  Group=paste(Age,Layer,sep=" "))%>%
    group_by(Group,PeakID)%>%summarise(mean=mean(value))%>%
    pivot_wider(names_from = Group,values_from=mean)%>%column_to_rownames("PeakID")



scale_counts_filt<-apply(log2(countmat_filt+0.1),1,FUN=function(x) (x-mean(x))/sd(x))

#gapstat<-clusGap(t(scale_counts_filt),FUNcluster = kmeans,K.max = 10,iter.max=30,nstart=100)
#km<-5 #for Introns
km<-2 #For PRomoters
#km<-6 #For IG
set.seed(404)
kmout<-kmeans(t(scale_counts_filt),centers=km,iter.max=30,nstart=100)

mat<-Heatmap(t(scale_counts_filt),
             cluster_columns = F,clustering_method_rows = "average",clustering_distance_rows = "pearson",show_row_names = F,
             col = colorRampPalette(c("blue2","white","red2"))(50),name="Normalized Expression",
             column_names_rot=45,heatmap_legend_param = list(direction="horizontal"),
             row_split = factor(as.numeric(kmout$cluster)),show_parent_dend_line = FALSE)
#levels=c(5,4,2,1,3) for intron ordering

#levels = c(4,2,1,3,6,5) for IG ordering

#Run to see heatmap and perform clustering
kmat<-draw(mat,heatmap_legend_side = "bottom")
```


```{r}
#retrieve results and group genes
klist<-row_order(kmat)
temp<-data.frame(RowID=do.call(c,klist),Cluster=rep(names(klist),sapply(klist,length)))

temp2<-as.data.frame(t(scale_counts_filt)[temp$RowID,])%>%mutate(Group=temp$Cluster)%>%rownames_to_column("ENSID")%>%
  pivot_longer(-c(ENSID,Group),names_to = "Sample")%>%
  group_by(Group,Sample)%>%mutate(avgexp=mean(value))

temp<-temp%>%mutate(PekeID=colnames(scale_counts_filt)[temp$RowID])

#Plot clusters
plotx<-temp2%>%ggplot(aes(x=Sample,y=avgexp,col=Group,group=Group))+geom_line()+geom_line(aes(y=value,group=ENSID),alpha=0.01)+
  stat_summary(mapping = aes(y=value),fun=mean,group=1,fun.min=function(x) mean(x)-sd(x),fun.max=function(x) mean(x)+sd(x),geom="errorbar",width=0.2)+
  labs(x="Group", y="Normalzied Expresion")+facet_wrap(~Group,ncol = 1)+theme_bw()+theme(axis.text.x = element_text(angle = 45,vjust=0.5))+
  labs(x=NULL,col="Cluster")+scale_color_brewer(palette = "Set1")
```


```{r}
library(TxDb.Mmusculus.UCSC.mm10.ensGene)
library(GenomicFeatures)
bfilt2<-getBM(attributes = c("ensembl_gene_id","ensembl_transcript_id","ucsc","external_gene_name","entrezgene_id","chromosome_name","transcript_biotype","transcript_gencode_basic"), mart = mart)
bfilt2<-bfilt2%>%filter(chromosome_name%in%c(as.character(1:19),"X","Y"),transcript_biotype%in%c("protein_coding"),transcript_gencode_basic=="GENCODE basic")

ensPromo<-promoters(TxDb.Mmusculus.UCSC.mm10.ensGene,upstream=2000,downstream=2000)
ensPromo_down<-ensPromo[ensPromo$tx_name%in%bfilt2$ensembl_transcript_id[bfilt2$ensembl_gene_id%in%L46downshared$ensembl_gene_id]]
ensPromo_up<-ensPromo[ensPromo$tx_name%in%bfilt2$ensembl_transcript_id[bfilt2$ensembl_gene_id%in%L46upshared$ensembl_gene_id]]
ensPromo_none<-ensPromo[ensPromo$tx_name%in%bfilt2$ensembl_transcript_id[bfilt2$ensembl_gene_id%in%cntset$ensembl_gene_id]]

TSSlarge<-TSSfrombmart(bmart2,extend = c(2000,2000),reduce=F)
RKPromo_down<-TSSlarge[TSSlarge$mcols.GeneID%in%L46downshared$ensembl_gene_id]
RKPromo_up<-TSSlarge[TSSlarge$mcols.GeneID%in%L46upshared$ensembl_gene_id]
RKPromo_none<-TSSlarge[TSSlarge$mcols.GeneID%in%cntset$ensembl_gene_id]
```

Get Regions overlapping peaks containing KLF motifs
```{r}
#load("~/Documents/DevATAC/p2p30PeaksSN_JASPAR2022CORE_FIMO.rda")
load("~/Documents/DevATAC/p2p30PeaksSN_JASPAR2022root_FIMO.rda")

peaksin2<-Propeaks
testout2<-sapply(listy,function(x) countOverlaps(peaksin2,x))
mcols(peaksin2)<-testout2
#KLFPromo<-peaksin2[peaksin2$KLF5_MA0599.1>0]
KLFPromo<-peaksin2[peaksin2$cluster_27>0]
CntPromo<-subsetByOverlaps(peaksin2,ProDAR,type="equal",invert = TRUE)
CntPromo<-subsetByOverlaps(CntPromo,KLFPromo,type="equal",invert=TRUE)

KLFPromo<-resize(KLFPromo,fix="center",width = 4000)
CntPromo<-resize(CntPromo,fix="center",width = 4000)

peaksin2<-allpeaks
testout2<-sapply(listy,function(x) countOverlaps(peaksin2,x))
mcols(peaksin2)<-testout2
#KLFAll<-peaksin2[peaksin2$KLF5_MA0599.1>0]
KLFAll<-peaksin2[peaksin2$cluster_27>0]
CntAll<-subsetByOverlaps(peaksin2,DARAnno,type="equal",invert = TRUE)
CntAll<-subsetByOverlaps(CntAll,KLFAll,type="equal",invert=TRUE)

KLFAll<-resize(KLFAll,fix="center",width = 4000)
CntAll<-resize(CntAll,fix="center",width = 4000)
```

Get top TSS Peaks (most likely DARs)
```{r}
TSSlarge<-TSSfrombmart(bmart2,extend = c(2000,2000),reduce=FALSE)
TSSnarrow<-TSSfrombmart(bmart2,extend=c(300,200),reduce = FALSE)
#Top Shared DEGs
testset<-rbind(L46downshared,L46upshared)
TSSset<-TSSnarrow[TSSnarrow$mcols.GeneID%in%testset$ensembl_gene_id]

#For KLF targets!
testset<-intergene%>%filter(ENSID%in%interbm$ensembl_gene_id[interbm$clustn%in%c(2,3)])
TSSset<-TSSnarrow[TSSnarrow$mcols.GeneID%in%testset$ENSID]

chk<-findOverlaps(TSSset,devcountGR)
TSSsub<-devcountGR[chk@to]
TSSsub$GeneID<-TSSset$mcols.GeneID[chk@from]
TSSsub$TxID<-TSSset$mcols.ENSID[chk@from]


#Find peaks overlapping TSS window with hightest coverage 
TSSsplit<-split(TSSsub,TSSsub$GeneID)
topTSS<-sapply(TSSsplit,FUN=function(x) x[which.max(apply(mcols(x)[,-c(9:10)],1,sum))])
topTSS<-unlist(GRangesList(topTSS))

topTSS$Change<-ifelse(topTSS$GeneID%in%L46upshared$ensembl_gene_id,"Up","Down")
topTSS<-split(topTSS,topTSS$Change)

#Unchanging Genes, Norm tests
cntset<-filter(L4_6df_filt,ENSID%in%names(sort(apply(rlg,1,var))[1:1000]))%>%filter(baseMeanL4>500)%>%
  left_join(bmart1[,1:2],by=c("ENSID"="external_gene_name"))
TSSset<-TSSnarrow[TSSnarrow$mcols.GeneID%in%cntset$ensembl_gene_id]

chk<-findOverlaps(TSSset,devcountGR)
TSSsub<-devcountGR[chk@to]
TSSsub$GeneID<-TSSset$mcols.GeneID[chk@from]
TSSsub$TxID<-TSSset$mcols.ENSID[chk@from]

#Find peaks overlapping TSS window with hightest coverage 
TSSsplit<-split(TSSsub,TSSsub$GeneID)
topTSS2<-sapply(TSSsplit,FUN=function(x) x[which.max(apply(mcols(x)[,-c(9:10)],1,sum))])
topTSS2<-unlist(GRangesList(topTSS2))
topTSS2$Change<-"None"

topTSS$None<-topTSS2

TSSout<-lapply(topTSS,function(x) ensPromo[ensPromo$tx_name%in%x$TxID])
#TSSout<-lapply(topTSS,function(x) TSSlarge[TSSlarge$mcols.ENSID%in%x$TxID])
```

Get Peaks around DEGs for coverage plots
```{r}
peaksets<-GRangesList(
ProDown=Propeaks[Propeaks$fixedGene%in%L46downshared$ensembl_gene_id],
ProUp=Propeaks[Propeaks$fixedGene%in%L46upshared$ensembl_gene_id],
ProCnt=Propeaks[Propeaks$fixedGene%in%cntset$ensembl_gene_id],
IntDown=Intpeaks[Intpeaks$fixedGene%in%L46downshared$ensembl_gene_id],
IntUp=Intpeaks[Intpeaks$fixedGene%in%L46upshared$ensembl_gene_id],
IntCnt=Intpeaks[Intpeaks$fixedGene%in%cntset$ensembl_gene_id],
IGDown=IGpeaks[IGpeaks$fixedGene%in%L46downshared$ensembl_gene_id],
IGUp=IGpeaks[IGpeaks$fixedGene%in%L46upshared$ensembl_gene_id],
IGCnt=IGpeaks[IGpeaks$fixedGene%in%cntset$ensembl_gene_id],
AllDown=AllPeaks[AllPeaks$fixedGene%in%L46downshared$ensembl_gene_id],
AllUp=AllPeaks[AllPeaks$fixedGene%in%L46upshared$ensembl_gene_id],
AllCnt=AllPeaks[AllPeaks$fixedGene%in%cntset$ensembl_gene_id])

peaksets<-resize(peaksets,fix="center",width = 4000)

peaksetsFULL<-GRangesList(
ProDown=PropeaksFULL[PropeaksFULL$fixedGene%in%L46downshared$ensembl_gene_id],
ProUp=PropeaksFULL[PropeaksFULL$fixedGene%in%L46upshared$ensembl_gene_id],
ProCnt=PropeaksFULL[PropeaksFULL$fixedGene%in%L46upshared$ensembl_gene_id],
IntDown=IntpeaksFULL[IntpeaksFULL$fixedGene%in%L46downshared$ensembl_gene_id],
IntUp=IntpeaksFULL[IntpeaksFULL$fixedGene%in%L46upshared$ensembl_gene_id],
IntCnt=IntpeaksFULL[IntpeaksFULL$fixedGene%in%cntset$ensembl_gene_id],
IGDown=IGpeaksFULL[IGpeaksFULL$fixedGene%in%L46downshared$ensembl_gene_id],
IGUp=IGpeaksFULL[IGpeaksFULL$fixedGene%in%L46upshared$ensembl_gene_id],
IGCnt=IGpeaksFULL[IGpeaksFULL$fixedGene%in%cntset$ensembl_gene_id],
AllDown=AllPeaksFULL[AllPeaksFULL$fixedGene%in%L46downshared$ensembl_gene_id],
AllUp=AllPeaksFULL[AllPeaksFULL$fixedGene%in%L46upshared$ensembl_gene_id],
AllCnt=AllPeaksFULL[AllPeaksFULL$fixedGene%in%cntset$ensembl_gene_id])

peaksetsFULL<-resize(peaksetsFULL,fix="center",width = 4000)
```

```{r}
#Promoters:
#Up Shared: Cluster 1; Down Shared: Cluster 2

#Introns:
#Up Shared: Cluster 1; Down Shared: 4 + 5

#Intragenic:
#Up Shared: Cluster 1; Down Shared: Cluster 3

upsharedpeak<-AllPeaks[as.numeric(names(kmout$cluster[kmout$cluster==1]))]
downsharedpeak<-AllPeaks[as.numeric(names(kmout$cluster[kmout$cluster==2]))]

uppeaknames<-unique(upsharedpeak$external_gene_name)
downpeaknames<-unique(downsharedpeak$external_gene_name)
geneuniv<-unique(L4_6df$ENSID)

GOup<- enrichGO(gene = uppeaknames, universe = geneuniv,
                OrgDb = org.Mm.eg.db,
                ont = "ALL",minGSSize = 25,
                pAdjustMethod = "BH",keyType = "SYMBOL",
                pvalueCutoff  = 0.05, qvalueCutoff = 0.05,
                readable = TRUE)
GOdown2<- enrichGO(gene = downpeaknames, universe = geneuniv2,
                OrgDb = org.Mm.eg.db,
                ont = "ALL",minGSSize = 25,
                pAdjustMethod = "BH",keyType = "SYMBOL",
                pvalueCutoff  = 0.05,qvalueCutoff = 0.05,
                readable = TRUE)

barplot(simplify(GOdown2,0.5),showCategory=10)+scale_y_discrete(position="right")

simplify(GOdown2,0.9)@result%>%as.data.frame()%>%mutate(yq=-log10(qvalue))%>%arrange(desc(yq))%>%head(10)%>%ggplot(aes(y=yq,x=reorder(Description,yq)))+geom_col(fill="mediumorchid",col="black")+coord_flip()+theme_classic()+labs(y=expression(-log[10](q.val)),x="")+theme(axis.text.y = element_text(size=16,face="bold"),text=element_text(size=18))+ scale_y_continuous(expand = c(0,0),limits = c(0,5))+scale_x_discrete(position = "top")
```

```{r}
#For Misc. Regulatory Regions
Regset<-PropeaksFULL
#Regset<-subsetByOverlaps(Regset, upsharedpeak,invert = T,minoverlap = 100)

ovtest<-findOverlaps(Regset,peaksin,type = "equal")
Regmotifs<-peaksin[ovtest@to]
Regmotifs$geneID<-Regset$fixedGene[ovtest@from]

#136 for clutser, 837 for all
testout_plus<-mcols(Regmotifs)[,-837]

listx<-list()
for(i in seq(ncol(testout_plus))){
  listx[[i]]<-testout_plus[,i]>0
}


temp<-overlapsAny(Regmotifs,downsharedpeak,minoverlap = 100)
contabs<-lapply(listx,FUN=contingency_mat)
names(contabs)<-names(listy)

Chitest_pos<-data.frame(Enrichment=sapply(contabs,FUN=function(x) (x[1,1]/sum(x[1,]))/(x[2,1]/sum(x[2,]))))%>%
  mutate(TFID=names(contabs),Pval=sapply(contabs,FUN=function(x) chisq.test(x)$p.value))%>%
  mutate(padj=p.adjust(Pval,"BH"))

Fishtest<-lapply(contabs,FUN=function(x) fisher.test(x))
Fishtest_pos<-data.frame(Enrichment=sapply(Fishtest,function(x) x$estimate),
                         Pval=sapply(Fishtest,function(x) x$p.value))%>%
                           mutate(TFID=names(contabs),padj=p.adjust(Pval,method="BH"))

volcanoooTF(Fishtest_pos,allpostext = T)
```

'Peaks first' Approach - measure DEGs associated with peaks in each category
```{r}
testkm<-genekm
peakkm<-kmout
DEGelig<-L4_6df_filt$ENSID
Peakset<-Propeaks
PeakDAR<-ProDAR

outdf<-data.frame()
for (i1 in c(1,seq(unique(peakkm$cluster))+1)){
  ix<-max(peakkm$cluster)
  if (i1==ix+1){peakset<-subsetByOverlaps(Peakset,PeakDAR,type="equal",invert = TRUE)$external_gene_name
  } else {
    peakset<-AllPeaks[as.numeric(names(which(peakkm$cluster==i1)))]$external_gene_name
  }
  n1<-length(peakset)
  n2<-sum(peakset%in%DEGelig)
  peakset_reduce<-unique(peakset)
  n3<-length(peakset_reduce)
  n4<-sum(peakset_reduce%in%DEGelig)
  tempdf<-data.frame(nPeaks=n1,nEligPeaks=n2,nUniPeaks=n3,nEligUniPeaks=n4)
  rm(n1,n2,n3,n4)
  for (i2 in seq(unique(testkm$cluster))) {
    geneset<-names(which(testkm$cluster==i2))
    n5<-sum(peakset%in%geneset)
    n6<-sum(peakset_reduce%in%geneset)
    id1<-paste0("nCluster_",as.character(i2))
    id2<-paste0("nUniCluster_",as.character(i2))
    tempdf[[id1]]<-n5
    tempdf[[id2]]<-n6
    rm(n5,n6,id1,id2)
  }
  outdf<-rbind(outdf,tempdf)
}

rownames(outdf)<-c(paste("Cluster",seq(unique(peakkm$cluster)),sep = "_"),"Unchanged_Peaks")

contabPeak<-function(dfout,onlyeligible=TRUE,mode=c("All","Binary")){
  if(onlyeligible){
    dfout<-dfout%>%dplyr::select(-c(nPeaks,nUniPeaks))
  } else {
    dfout<-dfout%>%dplyr::select(-c(nEligPeaks,nEligUniPeaks))
  }
  if(mode=="All"){
  dfout%>%dplyr::select(-contains("Uni"))%>%
    mutate(nNeggenes=.[,1]-apply(.[,-1],1,sum))%>%
      dplyr::select(-1)
  } else if(mode=="Binary"){
    dfout%>%dplyr::select(contains("Uni"))%>%
    mutate(nNeggenes=.[,1]-apply(.[,-1],1,sum))%>%
      dplyr::select(-1)
  }
}

temp<-contabPeak(outdf,mode="All")
names(temp)<-c("Shared Up","L6 Up","Shared Down","L6 Down","L4 Down","L4 Up","Not DEG")
temp<-temp[,c(1,3,7,5,6,2,4)]
chitemp<-chisq.test(temp)
corrplot(chitemp$residuals[,1:3],is.corr = F,method="color")
```

'Genes first' approach - how many DARs associated with genes in each category 
```{r}
library(chisq.posthoc.test)
library(corrplot)
#km67$cluster[km67$cluster==3]<-2
testkm<-km913_67
peakset<-Propeaks
peakkm<-kmout
#bgset<-L4_6df_filt%>%filter(!ENSID%in%names(testkm$cluster))
bgset<-K67df_10%>%filter(!external_gene_name%in%names(testkm$cluster))

outdf<-data.frame()
for (i1 in c(1,seq(unique(testkm$cluster))+1)){
  print(i1);
  if (i1==4){geneset<-bgset$external_gene_name} else {geneset<-names(which(testkm$cluster==i1))};
  n1<-length(geneset)
  n2<-sum(geneset%in%peakset$external_gene_name)
  n3<-sum(peakset$external_gene_name%in%geneset)
  tempdf<-data.frame(nGenes=n1,nGeneswPeaks=n2,nPeaks=n3)
  rm(n1,n2,n3)
  for (i2 in seq(unique(peakkm$cluster))) {
    testset<-AllPeaks[as.numeric(names(which(peakkm$cluster==i2)))]$external_gene_name
    n4<-sum(geneset%in%testset)
    n5<-sum(testset%in%geneset)
    id1<-paste0("nGenesCluster",as.character(i2))
    id2<-paste0("nPeaksCluster",as.character(i2))
    tempdf[[id1]]<-n4
    tempdf[[id2]]<-n5
    rm(n4,n5,id1,id2)
  }
  outdf<-rbind(outdf,tempdf)
}
rownames(outdf)<-c("Shared Up","L6 Up","Shared Down","L6 Down","L4 Down","L4 Up","Not DEG")
outdf<-outdf[c(1,3,7,5,6,2,4),]

rownames(outdf)<-c("Downregulated","Upregulated","Not DEG")
#rownames(outdf)<-c("Upregulated","Downregulated","Not DEG")
#rownames(outdf)<-c("Upregulated","Downregulated 1","Downregulated 2","Not DEG")
                   
contabGene<-function(dfout,level=c("Genes","Peaks")){
  if(level=="Genes"){
  dfout%>%dplyr::select(-nGenes)%>%
   dplyr::select(contains("Genes"))%>%
    mutate(nGenes_Neg=nGeneswPeaks-apply(.[,-1],1,sum))%>%
      dplyr::select(-nGeneswPeaks)
  } else if(level=="Peaks"){
    dfout%>%dplyr::select(-c(nGenes,nGeneswPeaks))%>%
    dplyr::select(contains("Peaks"))%>%
    mutate(nGenes_Neg=nPeaks-apply(.[,-1],1,sum))%>%
      dplyr::select(-nPeaks)
  }
}

temp<-contabGene(outdf,level="Peaks")
chitemp<-chisq.test(temp)
corrplot(chitemp$stdres[1:3,],is.corr = F,method="color",cl.cex=1.5,cl.align.text = 'l')
chisq.posthoc.test(temp)
#c(1,5,6,3,2,4,7)

names(temp)<-c("Gained","Lost","No Change")
temp[1:4,]%>%rownames_to_column('Cluster')%>%pivot_longer(cols=-Cluster)%>%mutate(Cluster=factor(Cluster,levels=c("Upregulated","Downregulated","Not DEG")))%>%ggplot(aes(x=Cluster,y=value,fill=name))+geom_bar(position="fill",stat="identity",col="black")+xlab("")+labs(fill="Promoter Accesibility Change")+scale_fill_manual(values = c("red","blue","grey70"))+theme_bw()+theme(axis.text=element_text(size = 15),axis.text.x=element_text(angle = 45,vjust =.6),axis.title = element_text(size=15),legend.position = "right")
```

```{r}
dcaGR_L6full<-dba.report(dca_out_L6,th=1)
seqlevelsStyle(dcaGR_L6full)<-"UCSC"
L6Pro<-subsetByOverlaps(dcaGR_L6full,Propeaks,type="equal")
L6Pro$ENSID<-Propeaks$fixedGene[subjectHits(findOverlaps(L6Pro,Propeaks,type="equal"))]
L6Pro$PeakReg=case_when(overlapsAny(L6Pro,dcaGR_L6[dcaGR_L6$Fold_L6>=1],type="equal") ~ "Gained",
                        overlapsAny(L6Pro,dcaGR_L6[dcaGR_L4$Fold_L6<=-1],type="equal") ~ "Lost",
                        TRUE ~ "No Change")

L6Pro$PeakReg=case_when(L6Pro$Fold>=1&L6Pro$FDR<=0.01 ~ "Gained",
                        L6Pro$Fold<=1&L6Pro$FDR<=0.01 ~ "Lost",
                        TRUE ~ "No Change")

dcaGR_L4full<-dba.report(dca_out_L4,th=1)
seqlevelsStyle(dcaGR_L4full)<-"UCSC"
L4Pro<-subsetByOverlaps(dcaGR_L4full,Propeaks,type="equal")
L4Pro$ENSID<-Propeaks$fixedGene[subjectHits(findOverlaps(L4Pro,Propeaks,type="equal"))]
L4Pro$PeakReg=case_when(overlapsAny(L4Pro,dcaGR_L4[dcaGR_L4$Fold_L4>=1],type="equal") ~ "Gained",
                        overlapsAny(L4Pro,dcaGR_L4[dcaGR_L4$Fold_L4<=-1],type="equal") ~ "Lost",
                        TRUE ~ "No Change")

L4Pro$PeakReg=case_when(L4Pro$Fold>=1&L4Pro$FDR<=0.01 ~ "Gained",
                        L4Pro$Fold<=1&L4Pro$FDR<=0.01 ~ "Lost",
                        TRUE ~ "No Change")

L4Comp<-L4df%>%mutate(GeneReg=case_when(log2FoldChange>=1&padj<=0.01 ~ "Upregulated", 
                                        log2FoldChange<=-1&padj<=0.01 ~ "Downregulated", 
                                        TRUE ~ "No Change"))%>%
  left_join(bmart1,by=c("ENSID"="external_gene_name"))%>%inner_join(as.data.frame(mcols(L4Pro)),by=c("ensembl_gene_id"="ENSID"))%>%
  dplyr::select(-starts_with('Conc'))%>%mutate(CellType="L4")

L6Comp<-L6df%>%mutate(GeneReg=case_when(log2FoldChange>=1&padj<=0.01 ~ "Upregulated", 
                                        log2FoldChange<=-1&padj<=0.01 ~ "Downregulated", 
                                        TRUE ~ "No Change"))%>%
  left_join(bmart1,by=c("ENSID"="external_gene_name"))%>%inner_join(as.data.frame(mcols(L6Pro)),by=c("ensembl_gene_id"="ENSID"))%>%
  dplyr::select(-starts_with('Conc'))%>%mutate(CellType="L6")

L46Comp<-rbind(L4Comp,L6Comp)%>%filter(!(GeneReg=="No Change"&PeakReg=="No Change"))

out<-L46Comp%>%ggplot(aes(x=log2FoldChange,y=Fold,shape=CellType))+
  geom_point(data=subset(L46Comp,GeneReg=="No Change"),color="grey70",alpha=0.25)+
  geom_pointdensity(data=subset(L46Comp,GeneReg=="Upregulated"),adjust=1)+
  scale_color_gradient(low = "indianred2",high = "darkred")+
  new_scale_color()+geom_pointdensity(data=subset(L46Comp,GeneReg=="Downregulated"),adjust=1)+
  scale_color_gradient(low="steelblue1",high="darkblue")+new_scale_color()+
    stat_smooth(aes(col=CellType),method=lm,fullrange = T,se = F)+
  stat_cor(aes(col=CellType,label = ..r.label..),label.x=-10,size=5)+
  scale_color_manual(values = c("purple","purple4"))+
       geom_hline(yintercept = c(-1,1),linetype="dashed")+geom_vline(xintercept = c(-1,1),linetype="dashed")+theme_bw()+
       labs(x="log2FoldChange(P2 v P30, RNA)",y="log2FoldChange(P2 v P30, ATAC)",shape="Cell Type")
```


```{r}
library(ggnewscale)
library(ggpointdensity)

#Compare P2 v P30 RNA-seq data with P2 v P30 ATAC-seq data
dcx_P2P30<-dba.contrast(dc_norm,contrast = c("Condition","P30","P2"),minMembers = 4)
dca_out_P2P30<-dba.analyze(dcx_P2P30)
dcaGR_P2P30<-dba.report(dca_out_P2P30,th=1)
seqlevelsStyle(dcaGR_P2P30)<-"UCSC"
P2P30Pro<-subsetByOverlaps(dcaGR_P2P30,Propeaks,type="equal")
P2P30Pro$ENSID<-Propeaks$fixedGene[subjectHits(findOverlaps(P2P30Pro,Propeaks,type="equal"))]
P2P30Pro$PeakReg=case_when(overlapsAny(P2P30Pro,upsharedpeak,type="equal") ~ "Gained", 
                           overlapsAny(P2P30Pro,downsharedpeak,type="equal") ~ "Lost",
                           TRUE ~ "No Change")

#(Devdf created in Figureplots.Rmd)
DevComp<-Devdf%>%inner_join(as.data.frame(mcols(P2P30Pro)),by=c("ensembl_gene_id"="ENSID"))%>%
  filter(ENSID%in%L4_6df_filt$ENSID)%>%mutate(GeneReg=case_when(ensembl_gene_id%in%L46upshared$ensembl_gene_id ~ "Upregulated",
                                                                ensembl_gene_id%in%L46downshared$ensembl_gene_id ~ "Downregulated", 
                                                                TRUE ~ "No Change"))%>%
  filter(!(GeneReg=="No Change"&PeakReg=="No Change"))%>%
  mutate(GenePeak=case_when(GeneReg=="Upregulated"&PeakReg=="Gained" ~ T,
                                                                 GeneReg=="Downregulated"&PeakReg=="Lost" ~ T,
                                                                 TRUE ~ F))%>%
  mutate(labid=ifelse(ENSID%in%topgenes,ENSID,""))

outplot<-DevComp%>%ggplot(aes(x=log2FoldChange,y=Fold,col=GeneReg,alpha=PeakReg))+geom_point()+
     scale_alpha_manual(values =c(1,0.5,0.1))+scale_color_manual(values=c("blue","grey70","red"))+
       geom_hline(yintercept = c(-1,1),linetype="dashed")+geom_vline(xintercept = c(-1,1),linetype="dashed")+theme_bw()+
       labs(x="log2FoldChange(P2 v P30, RNA)",y="log2FoldChange(P2 v P30, ATAC)",col="Gene Expression Change", alpha="Promoter Accesibility Change")+
       theme(text  = element_text(size=13),legend.position = "bottom")+
       geom_point(data=subset(DevComp,GenePeak==T),color="black",shape=21,alpha=1)+
  geom_text_repel(data=subset(DevComp,GenePeak==T),aes(label=ENSID),max.overlaps = 20)

outplot2<-DevComp%>%ggplot(aes(x=log2FoldChange,y=Fold))+
  geom_point(data=subset(DevComp,GeneReg=="No Change"),color="grey70",alpha=0.5)+
  geom_pointdensity(data=subset(DevComp,GeneReg=="Upregulated"),adjust=1)+
  scale_color_gradient(low = "indianred2",high = "darkred")+
  new_scale_color()+geom_pointdensity(data=subset(DevComp,GeneReg=="Downregulated"),adjust=1)+
  scale_color_gradient(low="steelblue1",high="darkblue")+
       geom_hline(yintercept = c(-1,1),linetype="dashed")+geom_vline(xintercept = c(-1,1),linetype="dashed")+theme_bw()+
       labs(x="log2FoldChange(P2 v P30, RNA)",y="log2FoldChange(P2 v P30, ATAC)")+
  new_scale_color()+
  geom_text_repel(data=subset(DevComp,labid!=""),aes(label=labid,col=GeneReg),max.overlaps = 30,box.padding = 0.5)+
  scale_color_manual(values=c("blue","red"))

outplot2+stat_smooth(col="grey25",method=lm,fullrange = T)+
  stat_cor(label.x=-15,label.y=5,size=5)

#For compairng with Klf9/13 data
K913Comp<-K913df_old%>%inner_join(as.data.frame(mcols(P2P30Pro)),by="ENSID")%>%
  mutate(GeneReg=case_when(ENSID%in%filter(K913df_filtold,log2FoldChange>=1)$ENSID ~ "Upregulated",
                                                                ENSID%in%filter(K913df_filtold,log2FoldChange<=-1)$ENSID ~ "Downregulated", 
                                                                TRUE ~ "No Change"))%>%
  filter(!(GeneReg=="No Change"&PeakReg=="No Change"))%>%
  mutate(GenePeak=case_when(GeneReg=="Upregulated"&PeakReg=="Lost" ~ T,
                                                                 GeneReg=="Downregulated"&PeakReg=="Gained" ~ T,
                                                                 TRUE ~ F))%>%
  filter(!external_gene_name%in%c("Klf9","Klf13"))%>%
  mutate(labid=ifelse(external_gene_name%in%KLFbestgenes,external_gene_name,""))

outplot<-K913Comp%>%
  ggplot(aes(x=log2FoldChange,y=Fold))+geom_point(aes(col=GeneReg,alpha=GeneReg))+
     scale_alpha_manual(values =c(1,0.25,1))+scale_color_manual(values=c("blue","grey70","red"))+
       geom_hline(yintercept = c(-1,1),linetype="dashed")+geom_vline(xintercept = c(-1,1),linetype="dashed")+theme_bw()+
       labs(x="log2FoldChange(Klf9/13 v Scr., RNA)",y="log2FoldChange(P2 v P30, ATAC)",
            col="Gene Expression Change", alpha="Promoter Accesibility Change")+
  theme(text  = element_text(size=13),legend.position = "bottom")+
  geom_text_repel(aes(label=labid,col=GeneReg),max.overlaps = 30,box.padding = 0.5,col="tomato4")+
    theme(axis.title =element_text(size=14),axis.text =element_text(size=13))

outplot+stat_smooth(col="grey25",method=lm,fullrange = T)+stat_cor(label.x=-4.5,label.y=5,size=5)

#For compairng with Klf6/7 data
K67comp<-K67df_10%>%inner_join(as.data.frame(mcols(P2P30Pro)),by="ENSID")%>%
  mutate(GeneReg=case_when(ENSID%in%filter(K67df10_filt,log2FoldChange>=1)$ENSID ~ "Upregulated",
                                                                ENSID%in%filter(K67df10_filt,log2FoldChange<=-1)$ENSID ~ "Downregulated", 
                                                                TRUE ~ "No Change"))%>%
  filter(!(GeneReg=="No Change"&PeakReg=="No Change"))%>%
  mutate(GenePeak=case_when(GeneReg=="Upregulated"&PeakReg=="Gained" ~ T,
                                                                 GeneReg=="Downregulated"&PeakReg=="Lost" ~ T,
                                                                 TRUE ~ F))%>%
  filter(!external_gene_name%in%c("Klf6","Klf7"))%>%
  mutate(labid=ifelse(external_gene_name%in%KLFbestgenes&GeneReg=="Downregulated",external_gene_name,""))

outplot<-K67comp%>%
  ggplot(aes(x=log2FoldChange,y=Fold))+geom_point(aes(col=GeneReg,alpha=GeneReg))+
     scale_alpha_manual(values =c(1,0.25,1))+scale_color_manual(values=c("blue","grey70","red"))+
       geom_hline(yintercept = c(-1,1),linetype="dashed")+geom_vline(xintercept = c(-1,1),linetype="dashed")+theme_bw()+
       labs(x="log2FoldChange(Klf6/7 v Scr., RNA)",y="log2FoldChange(P2 v P30, ATAC)",
            col="Gene Expression Change", alpha="Promoter Accesibility Change")+
  theme(text  = element_text(size=13),legend.position = "bottom")+
  geom_text_repel(aes(label=labid,col=GeneReg),max.overlaps = 30,box.padding = 0.5,col="blue4")+
    theme(axis.title =element_text(size=14),axis.text =element_text(size=13))

outplot+stat_smooth(col="grey25",method=lm,fullrange = T)+stat_cor(label.x=-4.5,label.y=5,size=5)
```

```{r}
ovhits<-findOverlaps(Propeaks,dcaGR_P2P30)
Propeaks$Fold<-dcaGR_P2P30$Fold[subjectHits(ovhits)]

temp<-data.frame(Target="Klf9/13",GeneID=filter(K913df_filt,log2FoldChange>=1)$ENSID)
temp<-rbind(temp,data.frame(Target="Klf6/7",GeneID=filter(K67df10_filt,log2FoldChange<=-1)$ENSID))
temp<-rbind(temp,data.frame(Target="Shared",GeneID=filter(intergene,log2FoldChange_Klf913>=1)$ENSID))
temp2<-data.frame(Target="Non-DEG",GeneID=filter(K67df_10,!ENSID%in%temp$GeneID)$ENSID)
temp<-rbind(temp,temp2[sample(seq(nrow(temp2)),2000),])

Profold<-temp%>%left_join(as.data.frame(mcols(Propeaks)),by=c("GeneID"="fixedGene"))%>%filter(!is.na(Fold))%>%
  mutate(Target=factor(Target,levels=c("Non-DEG","Klf9/13","Klf6/7","Shared")))
out<-Profold%>%ggplot(aes(x=Target,y=Fold,fill=Target))+geom_violin(draw_quantiles = c(0.25,0.5,0.75),scale="width")+
  scale_fill_manual(values = c("blue4","firebrick3","springgreen3","purple4"))+theme_classic()+
  labs(y=expression('Promoter log'[2]*'Fold Change'))+theme(axis.text= element_text(size=15),axis.title.y= element_text(size=15),legend.position = "none")
```

